---
title: "Study 2 — Behavioural Analysis"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r}
# ============================================================================
# study2_behavioural_analysis.Rmd — model-free + regression analyses, Study 2
#
# Aim:      Main behavioural analysis for Study 2 (risky vs sure choices in
#           gain and loss domains), organised to mirror the manuscript:
#           sample / preregistered exclusions and descriptives, complex-option
#           and risky-option choice t-tests, response-time analyses, the
#           Bayesian logistic mixed-effects regression (brms), subjective
#           complexity / cognitive-uncertainty ratings, cognitive-ability
#           correlations, high-EV-difference trial accuracy, the Appendix B
#           complex-role test, and the manuscript summary figures.
# Inputs:   ../data/study2/merged_data_study2.csv.gz (anonymised trial-level
#           pre-exclusion data; one row per jsPsych event)
# Outputs:  figure .png files (individual panels plus the manuscript
#           assemblies summary_analysis_1_update.png and
#           summary_analysis_2.png) and the brms model cache
#           m_brm_study2.rds, written to the working directory
#           (02_behavioural/)
# Exclusions applied here:
#   - Comprehension questions: participants with more than 8 total attempts
#     across the four comprehension ("design") questions are excluded.
#   - RT cleaning (test trials, practice excluded): trials with RT < 1 s or
#     RT > 30 s are removed; then, within participant x block, trials flagged
#     as beyond median +/- 3 SD are removed.
#   - Minimum-trials criterion: participants with fewer than 62 (= 124/2)
#     remaining test trials are excluded.
# Usage:    knit or run chunks from within 02_behavioural/ so the relative
#           input path resolves. The brms model is cached to
#           m_brm_study2.rds after the first fit; delete that file to force
#           a refit. No network access is required.
# ============================================================================
```

# Setup

Packages, data load, and consolidated helper functions (significance stars
and the shared plot themes) used throughout the report.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# Packages used throughout the analysis
library(data.table)
library(dplyr)
library(tidyverse)
library(jsonlite)
library(lubridate)
library(magrittr)
library(purrr)
library(knitr)
library(stringr)
library(broom)
library(cowplot)
library(brms)
library(parameters)
```

```{r load-data}
# Load anonymised trial-level data; fread() reads the .gz directly (via R.utils)
data <- fread("../data/study2/merged_data_study2.csv.gz")
```

```{r helpers}
# Shared helpers used throughout the notebook.

# p-value -> significance stars
p_to_stars <- function(p){
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ "ns"
  )
}

# Publication-style theme for the bar/jitter and correlation panels
theme_pub <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.title         = element_text(face = "bold"),
      axis.text          = element_text(color = "black"),
      panel.border       = element_rect(color = "black", fill = NA, linewidth = 0.5),
      plot.title         = element_text(face = "bold", hjust = 0.5),
      axis.line          = element_line(color = "black"),
      legend.position    = "none"
    )
}

# Variant used by the subjective-rating panels (legend on top, no axis line)
theme_pub_top <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.title         = element_text(face = "bold"),
      axis.text          = element_text(color = "black"),
      panel.border       = element_rect(color = "black", fill = NA, linewidth = 0.5),
      plot.title         = element_text(face = "bold", hjust = 0.5),
      legend.position    = "top"
    )
}

# Strip CSV-style quoting from JSON survey responses
clean_json <- function(x) {
  x <- gsub('^"|"$', "", x)             # strip outer quotes
  x <- gsub('""', '"', x, fixed = TRUE) # unescape CSV-style quotes
  x
}
```

# Preregistered exclusions

Participants answered four comprehension ("design") questions; anyone who
needed more than 8 total attempts to pass them is excluded. Response-time
cleaning then removes trials with RT < 1 s or > 30 s and, within
participant x block, trials beyond median +/- 3 SD; participants with fewer
than 62 (= 124/2) remaining test trials are excluded.

```{r comprehension-exclusion}
# Count comprehension-question attempts per participant and apply the
# preregistered exclusion (> 8 attempts across the four questions)
data_filtered <- data %>%
  group_by(prolific_pid) %>%
  mutate(attempts_to_pass_four_questions = sum(trial_type_label == "question", na.rm = TRUE)) %>%
  ungroup() %>%
  filter(attempts_to_pass_four_questions <= 8)
```

```{r rt-cleaning-absolute}
# Keep main test trials (practice excluded), convert RT from ms to seconds,
# and flag absolute outliers (RT < 1 s or RT > 30 s)
cleaned_RT_data <- data_filtered %>%
  filter(trial_type_label == 'test' & test_part != 'prac') %>%
  mutate(rt = as.numeric(rt_trial/1000))

cleaned_RT_data <- cleaned_RT_data %>%
  mutate(
    is_outlier1 = rt < 1 | rt > 30
  )
```

```{r rt-cleaning-median-sd}
# Within participant x block, flag trials beyond median +/- 3 SD and drop them
cleaned_RT_data <- cleaned_RT_data %>%
  filter(!is_outlier1 ) %>%
  group_by(prolific_pid, test_part) %>%
  mutate(
    median_rt = median(rt),
    sd_rt = sd(rt),
    is_outlier2 = rt < (median_rt - 3 * sd_rt) | rt > (median_rt + 3 * sd_rt)
  )

cleaned_RT_data <- cleaned_RT_data %>%
  filter(!is_outlier2)
```

```{r min-trials-criterion}
# Minimum-trials criterion: keep participants with at least 62 (= 124/2)
# clean test trials
trial_counts <- cleaned_RT_data %>%
  group_by(prolific_pid) %>%
  dplyr::summarize(trials = n())


participants_to_keep <- trial_counts %>%
  filter(trials >= 124/2) %>%
  select(prolific_pid)


cleaned_RT_data <- cleaned_RT_data %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid)
```

```{r exclusion-flow-table}
# Exclusion flow (150 recruited -> 10 comprehension-excluded ->
# 5 RT-excluded -> N = 135)
n_recruited  <- dplyr::n_distinct(data$prolific_pid)
n_after_comp <- dplyr::n_distinct(data_filtered$prolific_pid)
n_final      <- nrow(participants_to_keep)

tibble(
  Stage = c(
    "Recruited",
    "Excluded: comprehension questions (> 8 attempts)",
    "Excluded: response-time criteria",
    "Final sample"
  ),
  N = c(
    n_recruited,
    n_recruited - n_after_comp,
    n_after_comp - n_final,
    n_final
  )
) %>%
  kable(caption = "Exclusion flow")
```

```{r analysis-dataset}
# Analysis dataset used by all choice analyses below.
# Recode responses so real_response reflects the option actually chosen
# (screen sides were counterbalanced via risk_index), then code whether the
# complex option was chosen on each trial
df_test1 <- cleaned_RT_data %>%
  filter(trial_type_label == "test" & skew != "catch") %>%
  mutate(real_response = case_when(
    risk_index == 1 & response == "f" ~ "f",
    risk_index == 1 & response == "j" ~ "j",
    risk_index == -1 & response == "f" ~ "j",
    risk_index == -1 & response == "j" ~ "f"
  )) %>%
  mutate(
    complex_choice = case_when(
      complexity_option == "A" & real_response == "f" ~ 1L,
      complexity_option == "B" & real_response == "j" ~ 1L,
      TRUE ~ 0L
    )
  )
```

# Sample descriptives

In total, 10 participants were excluded based on the preregistered
design-question accuracy criterion, and an additional 5 participants were
excluded due to response-time criteria. The final sample for analysis
consisted of 135 participants (mean age = 39.32 years, SD = 10.19,
range = 19-60), of whom 44.4% were female (n = 60) and 55.6% were male
(n = 75).

## Demographics

```{r demographics-parse}
# Parse demographics (age, gender, education) and free-text feedback from
# the JSON-encoded survey responses
demo_table <- data %>%
  filter(trial_type_label == 'demo')

demo_clean <- demo_table %>%
  mutate(
    parsed = map(response, ~{
      txt <- clean_json(.x)
      tryCatch(fromJSON(txt), error = function(e) list())
    }),
    Age       = map_chr(parsed, ~pluck(.x, "Age",       .default = NA_character_)),
    Gender    = map_chr(parsed, ~pluck(.x, "Q0",        .default = NA_character_)),
    Education = map_chr(parsed, ~pluck(.x, "Q1",        .default = NA_character_)),
    Feedback  = map_chr(parsed, ~pluck(.x, "feedback",  .default = NA_character_))
  ) %>%
  select(prolific_pid, Age, Gender, Education, Feedback)

# make Age numeric
demo_clean <- demo_clean %>%
  mutate(Age = suppressWarnings(as.numeric(Age)))

# one row per participant (if duplicated, keep first non-NA)
demo_final <- demo_clean %>%
  group_by(prolific_pid) %>%
  summarise(
    Age       = suppressWarnings(as.numeric(na.omit(Age)[1])),
    Gender    = na.omit(Gender)[1]    %||% NA_character_,
    Education = na.omit(Education)[1] %||% NA_character_,
    Feedback  = na.omit(Feedback)[1]  %||% NA_character_,
    .groups = "drop"
  )
```

```{r demographics-summary}
# Sample demographics after all exclusions
demo_final_filtered <- demo_final %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid)

demo_summary <- demo_final_filtered %>%
  summarise(
    n = n(),
    mean_age = mean(Age, na.rm = TRUE),
    sd_age = sd(Age, na.rm = TRUE),
    min_age = min(Age, na.rm = TRUE),
    max_age = max(Age, na.rm = TRUE),
    female_n = sum(Gender == "Female", na.rm = TRUE),
    male_n = sum(Gender == "Male", na.rm = TRUE),
    other_n = sum(!(Gender %in% c("Male", "Female"))),
    female_pct = round(100 * female_n / n, 1),
    male_pct   = round(100 * male_n / n, 1),
    other_pct  = round(100 * other_n / n, 1)
  )

demo_summary %>%
  kable(digits = 2, caption = "Demographics of the final sample (N = 135)")
```

## Comprehension-question attempts

```{r comprehension-attempts}
cleaned_RT_data %>%
  ungroup() %>%
  summarise(mean_attempt = mean(attempts_to_pass_four_questions)) %>%
  kable(digits = 2,
        caption = "Mean comprehension-question attempts (final sample)")
```

## Median response times by domain

Overall median RTs per domain and the paired comparison of per-participant
median RTs (losses vs gains). This domain RT comparison is reported in the
Discussion (alternative explanations).

```{r rt-by-domain}
df_rt_gain_loss_diff <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by(subject, test_part) %>%
  summarise(median_rt = median(rt), .groups = "drop")

df_rt_gain_loss_diff_median <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by( test_part) %>%
  summarise(median_rt = median(rt),
            sd_rt = sd(rt), .groups = "drop")

df_rt_gain_loss_diff_median %>%
  kable(digits = 2, caption = "Median RT (s) per domain")

df_rt_gain_loss_wide <- df_rt_gain_loss_diff %>%
  tidyr::pivot_wider(
    names_from  = test_part,
    values_from = median_rt
  )

t_rt_domain <- t.test(
  df_rt_gain_loss_wide$Loss,
  df_rt_gain_loss_wide$Gain,
  paired = TRUE
)

tidy(t_rt_domain) %>%
  kable(digits = 3,
        caption = "Paired t-test: per-participant median RT, Loss vs Gain")
```

## Bonus payments

```{r bonus-summary}
bonus_summary <- data %>%
  filter(!is.na(total_bonus)) %>%  # Keep rows where total_bonus is not NA
  distinct(prolific_pid, total_bonus)

bonus_summary %>%
  summarise(
    n = n(),
    mean_bonus_gbp = mean(total_bonus),
    min_bonus_gbp  = min(total_bonus),
    max_bonus_gbp  = max(total_bonus)
  ) %>%
  kable(digits = 2, caption = "Bonus payments (GBP)")
```

## Completion time

```{r completion-time}
# Completion time per participant = final cumulative time stamp of their
# session (time_elapsed is cumulative within a session, in ms)
completion <- data %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid) %>%
  group_by(prolific_pid) %>%
  summarise(minutes = max(time_elapsed, na.rm = TRUE) / 1000 / 60,
            .groups = "drop")

kable(
  tibble(`Median completion time (min)` = median(completion$minutes)),
  digits = 1,
  caption = "Completion time."
)
```

# Complex-option choice

Two-sided one-sample t-tests comparing the mean choice proportion of complex
options against the 50% benchmark, separately for the gain and loss
conditions, plus the paired comparison between domains.

```{r complex-choice-tests}
complex_choice_summary_individual <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by(prolific_pid, test_part) %>%
  summarise(
    n_trials = n(),
    n_complex = sum(complex_choice, na.rm = TRUE),
    prop_complex = mean(complex_choice, na.rm = TRUE),
    .groups = "drop"
  )

complex_choice_summary_group <-  complex_choice_summary_individual %>%
  group_by(test_part) %>%
  summarise(
    mean_complex_prop = mean(prop_complex, na.rm = TRUE),
    se = sd(prop_complex, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_complex_prop - 1.96 * se,
    upper = mean_complex_prop + 1.96 * se
  )

# ---------- t-tests ----------
t_gain <- t.test(
  complex_choice_summary_individual$prop_complex[
    complex_choice_summary_individual$test_part == "Gain"
  ],
  mu = 0.5
)
t_loss <- t.test(
  complex_choice_summary_individual$prop_complex[
    complex_choice_summary_individual$test_part == "Loss"
  ],
  mu = 0.5
)

complex_wide <- complex_choice_summary_individual %>%
  tidyr::pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = prop_complex
  )

t_gain_vs_loss <- t.test(
  complex_wide$Gain, complex_wide$Loss,
  paired = TRUE
)

complex_choice_summary_group %>%
  kable(digits = 4, caption = "Complex-option choice proportion by domain")

# tidy summary table
complex_t_summary <- dplyr::bind_rows(
  broom::tidy(t_gain)         %>% dplyr::mutate(test = "Gain vs 0.5"),
  broom::tidy(t_loss)         %>% dplyr::mutate(test = "Loss vs 0.5"),
  broom::tidy(t_gain_vs_loss) %>% dplyr::mutate(test = "Gain vs Loss (paired)")
)

complex_t_summary %>%
  kable(digits = 4, caption = "Complex-option choice t-tests")
```

```{r complex-choice-bf}
# Bayesian one-sample t-test vs 0.5 for the loss-domain complex-choice
# proportions (Appendix A), as in Study 1.
library(BayesFactor)
bf_loss <- ttestBF(
  complex_choice_summary_individual$prop_complex[
    complex_choice_summary_individual$test_part == "Loss"],
  mu = 0.5
)
kable(
  tibble(`BF10` = extractBF(bf_loss)$bf, `BF01` = 1 / extractBF(bf_loss)$bf),
  digits = 3,
  caption = "Bayes factors for the loss-domain complex-choice proportion vs 0.5 (default Cauchy prior, r = 0.707)."
)
```

```{r complex-choice-plot, fig.width=7, fig.height=5}
stars_gain <- p_to_stars(t_gain$p.value)
stars_loss <- p_to_stars(t_loss$p.value)
stars_pair <- p_to_stars(t_gain_vs_loss$p.value)

# ---------- plotting data ----------
plot_df_points <- complex_choice_summary_individual
plot_df_bars   <- complex_choice_summary_group

plot_df_bars$test_part   <- factor(plot_df_bars$test_part,
                                   levels = c("Gain","Loss"),
                                   labels = c("Gains", "Losses"))
plot_df_points$test_part <- factor(plot_df_points$test_part,
                                   levels = c("Gain","Loss"),
                                   labels = c("Gains", "Losses"))

# ---- fixed positions for stars & bracket ----
y_gain_star <- 0.85
y_loss_star <- 0.85
bracket_y   <- 0.93   # ends of the bracket
label_y     <- 0.95  # text for Gain vs Loss

# ---------- plot ----------
plot_complex_choice <- ggplot() +
  geom_jitter(
    data = plot_df_points,
    aes(x = test_part, y = prop_complex),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  geom_col(
    data = plot_df_bars,
    aes(x = test_part, y = mean_complex_prop),
    width = 0.55,
    fill = "grey60",
    color = "black",
    alpha = 0.2
  ) +
  geom_errorbar(
    data = plot_df_bars,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.15,
    linewidth = 1.3

  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  # one-sample stars (vs .5), now clearly above dots
  geom_text(aes(x = "Gains",  y = y_gain_star, label = stars_gain), vjust = 0, size = 6, fontface = "bold") +
  geom_text(aes(x = "Losses", y = y_loss_star, label = stars_loss), vjust = 0, size = 6, fontface = "bold") +
  # bracket + star for Gain vs Loss
  geom_segment(aes(x = 1, xend = 1, y = bracket_y - 0.01, yend = bracket_y)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y - 0.01, yend = bracket_y)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y,       yend = bracket_y)) +
  geom_text(aes(x = 1.5, y = label_y, label = stars_pair), vjust = 0, size = 6, fontface = "bold") +
  scale_y_continuous(
    limits = c(0, 1),              # allow a bit of space above 1
    breaks = seq(0, 1, by = 0.25),
    expand = c(0, 0)
  ) +
  labs(
    x = "Domain",
    y = "Proportion of complex-option choices",
    title = "Complex-option choices in gain and loss domains"
  ) +
  theme_pub()

plot_complex_choice

ggsave(
  filename = "s2_plot_complex_choice.png",
  plot = plot_complex_choice,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# Risky-option choice

Two-sided one-sample t-tests comparing the mean choice proportion of risky
options against the 50% benchmark, separately for the gain and loss
conditions, plus the paired comparison between domains.

```{r risky-choice-tests}
risk_choice_summary <- df_test1 %>%
  group_by(prolific_pid, test_part) %>%
  summarise(
    n_trials = n(),
    n_risky = sum(real_response == "f", na.rm = TRUE),
    prop_risky = mean(real_response == "f", na.rm = TRUE),
    .groups = "drop"
  )

risky_choice_summary_group <-  risk_choice_summary %>%
  group_by(test_part) %>%
  summarise(
    mean_risky_prop = mean(prop_risky, na.rm = TRUE),
    se = sd(prop_risky, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_risky_prop - 1.96 * se,
    upper = mean_risky_prop + 1.96 * se
  )

# --- One-sample (vs .5) t-tests for prop_risky ---
t_risk_gain <- t.test(
  risk_choice_summary$prop_risky[risk_choice_summary$test_part == "Gain"],
  mu = 0.5, alternative = "two.sided"
)

t_risk_loss <- t.test(
  risk_choice_summary$prop_risky[risk_choice_summary$test_part == "Loss"],
  mu = 0.5, alternative = "two.sided"
)

# --- Paired t-test Gain vs Loss ---
risk_wide <- risk_choice_summary %>%
  tidyr::pivot_wider(
    id_cols   = prolific_pid,
    names_from = test_part,
    values_from = prop_risky
  )

t_risk_gain_vs_loss <- t.test(
  risk_wide$Gain, risk_wide$Loss,
  paired = TRUE, alternative = "two.sided"
)

risky_choice_summary_group %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  kable(digits = 4, caption = "Risky-option choice proportion by domain")

# tidy summary table
risk_t_summary <- dplyr::bind_rows(
  broom::tidy(t_risk_gain)         %>% dplyr::mutate(test = "Gain vs 0.5"),
  broom::tidy(t_risk_loss)         %>% dplyr::mutate(test = "Loss vs 0.5"),
  broom::tidy(t_risk_gain_vs_loss) %>% dplyr::mutate(test = "Gain vs Loss (paired)")
)

risk_t_summary %>%
  kable(digits = 4, caption = "Risky-option choice t-tests")
```

```{r risky-choice-plot, fig.width=7, fig.height=5}
stars_gain_r <- p_to_stars(t_risk_gain$p.value)
stars_loss_r <- p_to_stars(t_risk_loss$p.value)
stars_pair_r <- p_to_stars(t_risk_gain_vs_loss$p.value)

# ---------- data for plotting ----------
plot_pts_r <- risk_choice_summary %>%
  mutate(test_part = factor(test_part,
                            levels = c("Gain","Loss"),
                            labels = c("Gains","Losses")))

plot_bars_r <- risky_choice_summary_group %>%
  mutate(test_part = factor(test_part,
                            levels = c("Gain","Loss"),
                            labels = c("Gains","Losses")))

# fixed positions for stars & bracket (all above the dots/bars)
y_gain_star_r <- 1
y_loss_star_r <- 1
bracket_y_r   <- 1.05
label_y_r     <- 1.07

# ---------- plot ----------
plot_risky_choice <- ggplot() +
  # individual participants
  geom_jitter(
    data = plot_pts_r,
    aes(x = test_part, y = prop_risky),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  # group mean bars (semi-transparent)
  geom_col(
    data = plot_bars_r,
    aes(x = test_part, y = mean_risky_prop),
    width  = 0.55,
    fill   = "grey60",
    color  = "black",
    alpha  = 0.45
  ) +
  # 95% CI
  geom_errorbar(
    data = plot_bars_r,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.12,
    linewidth = 1.2
  ) +
  # reference line at 0.5
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  # one-sample stars above each bar (vs .5)

geom_text(
  aes(x = "Gains",  y = y_gain_star_r, label = stars_gain_r),
  vjust = 0, size = 6, fontface = "bold"
) +
geom_text(
  aes(x = "Losses", y = y_loss_star_r, label = stars_loss_r),
  vjust = 0, size = 6, fontface = "bold"
) +

# bracket + star for Gain vs Loss (paired)
geom_segment(aes(x = 1, xend = 1, y = 1.07, yend = 1.07)) +
geom_segment(aes(x = 2, xend = 2, y = 1.07, yend = 1.07)) +
geom_segment(aes(x = 1, xend = 2, y = 1.07, yend = 1.07)) +
geom_text(
  aes(x = 1.5, y = label_y_r, label = stars_pair_r),
  vjust = 0, size = 6, fontface = "bold"
) +
  scale_y_continuous(
    limits = c(0, 1.15),
    breaks = seq(0, 1, by = 0.25),
    expand = c(0, 0)
  ) +
  labs(
    x = "Domain",
    y = "Proportion of risky choices",
    title = "Risky choices in gain and loss domains"
  ) +
  theme_pub()

plot_risky_choice

ggsave(
  filename = "s2_plot_risky_choice.png",
  plot = plot_risky_choice,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# Response times

## Median RT: complex vs simple choices

Per-participant median RT difference (complex-chosen minus simple-chosen
trials), tested against 0 within each domain and compared between domains.

```{r rt-complex-simple-tests}
# ---- per participant × condition × choice (complex vs simple) ----
rt_individual <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by(prolific_pid, test_part, complex_choice) %>%
  summarise(
    median_rt = median(rt, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

# complex_choice: 1 = complex chosen, 0 = simple chosen
rt_wide <- rt_individual %>%
  mutate(choice_label = ifelse(complex_choice == 1, "complex", "simple")) %>%
  select(prolific_pid, test_part, choice_label, median_rt) %>%
  pivot_wider(
    id_cols   = c(prolific_pid, test_part),
    names_from = choice_label,
    values_from = median_rt,
    names_prefix = "median_rt_"
  ) %>%
  # keep only rows with both complex & simple
  filter(!is.na(median_rt_complex), !is.na(median_rt_simple))


# ---- per-participant RT difference (complex − simple) ----
rt_diff <- rt_wide %>%
  transmute(
    prolific_pid,
    test_part,
    rt_diff = median_rt_complex - median_rt_simple
  ) %>%
  mutate(
    test_part = factor(test_part, levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses"))
  )

# ---- group summary for bars + 95% CI ----
rt_diff_summary <- rt_diff %>%
  group_by(test_part) %>%
  summarise(
    mean_diff = mean(rt_diff, na.rm = TRUE),
    se        = sd(rt_diff,  na.rm = TRUE) / sqrt(n()),
    lower     = mean_diff - 1.96 * se,
    upper     = mean_diff + 1.96 * se,
    .groups   = "drop"
  )

# ---- tests: one-sample vs 0 (within domain) ----
t_gain_median <- t.test(
  rt_diff$rt_diff[rt_diff$test_part == "Gains"],
  mu = 0
)
t_loss_median <- t.test(
  rt_diff$rt_diff[rt_diff$test_part == "Losses"],
  mu = 0
)

# ---- paired Gain vs Loss (within-subjects) ----
rt_diff_wide <- rt_diff %>%
  pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = rt_diff
  )

t_pair_median <- t.test(
  rt_diff_wide$Gains,
  rt_diff_wide$Losses,
  paired = TRUE
)

# ---- tidy summary table ----
rt_tests_summary_median <- bind_rows(
  tidy(t_gain_median)  %>% mutate(test = "Gains vs 0",  metric = "Median RT (complex − simple)"),
  tidy(t_loss_median)  %>% mutate(test = "Losses vs 0", metric = "Median RT (complex − simple)"),
  tidy(t_pair_median)  %>% mutate(test = "Gains vs Losses (paired)", metric = "Median RT (complex − simple)")
) %>%
  select(metric, test, estimate, statistic, parameter, p.value, conf.low, conf.high)

rt_tests_summary_median %>%
  kable(digits = 4, caption = "Median RT differences: complex − simple")
```

```{r rt-complex-simple-plot, fig.width=7, fig.height=5}
stars_gain   <- p_to_stars(t_gain_median$p.value)
stars_loss   <- p_to_stars(t_loss_median$p.value)
stars_pair   <- p_to_stars(t_pair_median$p.value)

# ---- positions for stars and bracket ----
rng       <- range(rt_diff_summary$upper, rt_diff_summary$lower, na.rm = TRUE)
span      <- diff(rng)
span      <- ifelse(span == 0, 0.1, span)  # fallback

y_gain_star <- 4
y_loss_star <- 4

bracket_y   <- 4.5
label_y     <- 4.3

# ---- plot ----
plot_rt_median_diff <- ggplot() +
  # individual participant differences
  geom_jitter(
    data  = rt_diff,
    aes(x = test_part, y = rt_diff),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  # group mean bars (semi-transparent)
  geom_col(
    data  = rt_diff_summary,
    aes(x = test_part, y = mean_diff),
    width = 0.55,
    fill  = "grey60",
    color = "black",
    alpha = 0.45
  ) +
  # 95% CI
  geom_errorbar(
    data = rt_diff_summary,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.12,
    linewidth = 1.2
  ) +
  # zero reference line
  geom_hline(yintercept = 0, linetype = "dashed") +
  # stars for one-sample tests
  geom_text(
    aes(x = "Gains",  y = y_gain_star, label = stars_gain),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  geom_text(
    aes(x = "Losses", y = y_loss_star, label = stars_loss),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  # bracket + star for Gain vs Loss
  geom_segment(aes(x = 1, xend = 1, y = bracket_y - 0.01 * span, yend = bracket_y)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y - 0.01 * span, yend = bracket_y)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y,               yend = bracket_y)) +
  geom_text(
    aes(x = 1.5, y = 4.6, label = stars_pair),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  labs(
    x = "Domain",
    y = "Median RT difference (complex − simple, s)",
    title = "Median RT differences for complex vs simple choices"
  ) +
  coord_cartesian(
    ylim = c(-3, 5
    )
  ) +
  theme_pub()

plot_rt_median_diff

ggsave(
  filename = "s2_plot_rt_median_diff.png",
  plot = plot_rt_median_diff,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

## Median RT: risky vs sure choices

```{r rt-risky-sure-data}
# Recode choice: risky = 1, safe = 0
df_rt_risky <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  mutate(
    risky_choice = ifelse(real_response == "f", 1,
                          ifelse(real_response == "j", 0, NA))
  ) %>%
  filter(!is.na(risky_choice))

# ---- summarise median RT per participant × domain × risky/safe ----
rt_individual_risky <- df_rt_risky %>%
  group_by(prolific_pid, test_part, risky_choice) %>%
  summarise(
    median_rt = median(rt, na.rm = TRUE),
    .groups = "drop"
  )

# ---- wide format (median_rt_risky vs median_rt_safe) ----
rt_wide_risky <- rt_individual_risky %>%
  mutate(choice_label = ifelse(risky_choice == 1, "risky", "safe")) %>%
  select(prolific_pid, test_part, choice_label, median_rt) %>%
  pivot_wider(
    id_cols = c(prolific_pid, test_part),
    names_from = choice_label,
    values_from = median_rt,
    names_prefix = "median_rt_"
  ) %>%
  filter(!is.na(median_rt_risky), !is.na(median_rt_safe))

# ---- compute difference (risky − safe) ----
rt_diff_risky <- rt_wide_risky %>%
  transmute(
    prolific_pid,
    test_part = factor(test_part,
                       levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses")),
    rt_diff = median_rt_risky - median_rt_safe
  )
```

```{r rt-risky-sure-tests}
# ---- one-sample t-tests vs 0 ----
t_gain_risky <- t.test(
  rt_diff_risky$rt_diff[rt_diff_risky$test_part == "Gains"],
  mu = 0
)

t_loss_risky <- t.test(
  rt_diff_risky$rt_diff[rt_diff_risky$test_part == "Losses"],
  mu = 0
)

# ---- paired t-test between domains ----
rt_diff_risky_wide <- rt_diff_risky %>%
  pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = rt_diff
  )

t_pair_risky <- t.test(
  rt_diff_risky_wide$Gains,
  rt_diff_risky_wide$Losses,
  paired = TRUE
)

# ---- tidy table ----
rt_tests_summary_risky <- bind_rows(
  tidy(t_gain_risky) %>% mutate(test = "Gains vs 0"),
  tidy(t_loss_risky) %>% mutate(test = "Losses vs 0"),
  tidy(t_pair_risky) %>% mutate(test = "Gains vs Losses (paired)")
) %>%
  mutate(metric = "Median RT (risky − safe)") %>%
  select(metric, test, estimate, statistic, parameter, p.value, conf.low, conf.high)

rt_tests_summary_risky %>%
  kable(digits = 4, caption = "Median RT differences: risky − sure")
```

```{r rt-risky-sure-plot, fig.width=7, fig.height=5}
# ---- summary for bars ----
rt_diff_risky_summary <- rt_diff_risky %>%
  group_by(test_part) %>%
  summarise(
    mean_diff = mean(rt_diff, na.rm = TRUE),
    se        = sd(rt_diff, na.rm = TRUE) / sqrt(n()),
    lower     = mean_diff - 1.96 * se,
    upper     = mean_diff + 1.96 * se,
    .groups = "drop"
  )

# ---- stars ----
stars_gain <- p_to_stars(t_gain_risky$p.value)
stars_loss <- p_to_stars(t_loss_risky$p.value)
stars_pair <- p_to_stars(t_pair_risky$p.value)

# vertical spacing
rng  <- range(rt_diff_risky_summary$lower, rt_diff_risky_summary$upper, na.rm = TRUE)
span <- diff(rng); if (span == 0) span <- 0.1

y_gain_star <- 4.1
y_loss_star <- 4.1

bracket_y <- 4.7
label_y   <- 4.9

pair_label <- stars_pair

# ---- plot ----
plot_rt_risky <- ggplot() +
  geom_jitter(
    data = rt_diff_risky,
    aes(x = test_part, y = rt_diff),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  geom_col(
    data = rt_diff_risky_summary,
    aes(x = test_part, y = mean_diff),
    width = 0.55,
    fill  = "grey60",
    color = "black",
    alpha = 0.45
  ) +
  geom_errorbar(
    data = rt_diff_risky_summary,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.12,
    linewidth = 1
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +

  geom_text(aes(x = "Gains",  y = y_gain_star, label = stars_gain),
            vjust = 0, size = 6, fontface = "bold") +
  geom_text(aes(x = "Losses", y = y_loss_star, label = stars_loss),
            vjust = 0, size = 6, fontface = "bold") +

  geom_segment(aes(x = 1, xend = 1, y = bracket_y - 0.01*span, yend = bracket_y)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y - 0.01*span, yend = bracket_y)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y,               yend = bracket_y)) +
  geom_text(aes(x = 1.5, y = label_y, label = pair_label),
            vjust = 0, size = 5, fontface = "bold") +

  labs(
    x = "Domain",
    y = "Median RT difference (risky − safe, s)",
    title = "Median RT differences for risky vs. safe choices"
  ) +
  coord_cartesian(
    ylim = c(-5,5)
  ) +
  theme_pub()

plot_rt_risky

ggsave(
  filename = "s2_plot_rt_risky.png",
  plot = plot_rt_risky,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# Bayesian mixed-effects regression

Bayesian logistic mixed-effects model of choosing the left option, with EV
and skewness difference covariates and the full complexity x risk (SD) x
domain interaction (the model reported in Table 2 of the paper).

```{r regression-covariates}
# Per-option payoff moments (mean, SD, skewness) and their A - B
# differences, used as regression covariates
df_test1 <- df_test1 %>%
  rowwise() %>%
  mutate(
    # --- Option A ---
    mean_A = probA1 * payA1 + probA2 * payA2,
    var_A  = probA1 * (payA1 - mean_A)^2 + probA2 * (payA2 - mean_A)^2,
    sd_A   = sqrt(var_A),
    skew_A = ifelse(sd_A > 0,
                    (probA1 * (payA1 - mean_A)^3 + probA2 * (payA2 - mean_A)^3) / (sd_A^3),
                    0),

    # --- Option B ---
    mean_B = probB1 * payB1 + probB2 * payB2,
    var_B  = probB1 * (payB1 - mean_B)^2 + probB2 * (payB2 - mean_B)^2,
    sd_B   = sqrt(var_B),
    skew_B = ifelse(sd_B > 0,
                    (probB1 * (payB1 - mean_B)^3 + probB2 * (payB2 - mean_B)^3) / (sd_B^3),
                    0),

    # --- Skewness difference ---
    skew_diff = skew_A - skew_B
  ) %>%
  ungroup()
```

```{r regression-fit}
# prep
df_test1_fix <- df_test1 %>%
  mutate(
    real_response_num = ifelse(response == "f", 1, 0),
    test_part      = factor(test_part, levels = c("Gain","Loss")),  # domain
    complexity_left = ifelse(risk_index == 1, ifelse(complexity_option == "A", 1L, 0L), ifelse(complexity_option == "A", 0L, 1L)),
    SD_diff_numeric = as.numeric(SD_diff_level)*as.numeric(risk_index),
    EVD_numeric     = as.numeric(EV_diff_level)*as.numeric(risk_index),
    EVD_z = scale(EVD_numeric, center = TRUE, scale = TRUE)[,1],
    SD_diff_z = scale(SD_diff_numeric, center = TRUE, scale = TRUE)[,1],
    skew_diff_z = scale(skew_diff*as.numeric(risk_index), center = TRUE, scale = TRUE)[,1]
  )

# Bayesian logistic mixed-effects model of choosing the left option (brms).
# Long-running (4 chains x 4000 iterations; can take tens of minutes).
# The fitted model is cached to an .rds file: caching only avoids refitting
# on re-knit and does not affect the estimates (model formula, data, and
# sampler settings are unchanged). Delete m_brm_study2.rds to force a refit.
brm_cache_file <- "m_brm_study2.rds"

if (file.exists(brm_cache_file)) {
  m_all_rs_brm <- readRDS(brm_cache_file)
} else {
  m_all_rs_brm <- brm(
    real_response_num ~
      EVD_z + skew_diff_z +
      complexity_left * SD_diff_z * test_part +
      (1 | prolific_pid),
    data = df_test1_fix,
    family = bernoulli(link = "logit"),
    chains = 4, cores = 4, iter = 4000,
    control = list(adapt_delta = 0.95, max_treedepth = 12)
  )
  saveRDS(m_all_rs_brm, brm_cache_file)
}
```

```{r regression-table}
# effects = "all" also reports the random-effect SDs (the model includes
# random slopes for complexity, risk, and domain, not only random intercepts)
model_parameters(m_all_rs_brm, ci = 0.95, effects = "all") %>%
  kable(digits = 3,
        caption = "Bayesian logistic mixed-effects regression (Table 2)")
```

# Subjective ratings

After the choice blocks, participants rated a subset of gambles on perceived
complexity and value certainty (1-7 Likert scales).

## Cognitive uncertainty (certainty) ratings

```{r certainty-rating-data}
# ============================================================
# Prepare data: trial-level cognitive-uncertainty (confidence) ratings by
# option complexity (simple/complex) and risky/sure status
# ============================================================

rating <- data_filtered %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid) %>%
  filter(option_complexity %in% c("complex", "simple")) %>%
  mutate(
    confidence_rating = confidence_rating + 1,
    option_risky_level = case_when(
      option_certainty == "sure" ~ "Sure (100%)",
      option_certainty == "uncertain" ~ "Risky (<100%)",
      TRUE ~ option_certainty
    ),
    option_risky_level = factor(option_risky_level,
                                levels = c("Risky (<100%)","Sure (100%)")),
    option_complexity  = factor(option_complexity,
                                levels = c("simple","complex"))
  )

# ============================================================
# Summary data for bars + CI
# ============================================================

rating_summary <- rating %>%
  group_by(test_part, option_complexity, option_risky_level) %>%
  summarise(
    mean_certainty = mean(confidence_rating),
    se = sd(confidence_rating)/sqrt(n()),
    lower = mean_certainty - 1.96*se,
    upper = mean_certainty + 1.96*se,
    .groups = "drop"
  )

# ============================================================
# RISKY vs SURE t-tests (per domain × complexity)
# ============================================================

t_tests_risky_sure <- rating %>%
  group_by(test_part, option_complexity, subject) %>%
  summarise(
    risky  = mean(confidence_rating[option_risky_level == "Risky (<100%)"]),
    sure   = mean(confidence_rating[option_risky_level == "Sure (100%)"]),
    .groups="drop"
  ) %>%
  group_by(test_part, option_complexity) %>%
  summarise(
    t_res = list(t.test(risky, sure, paired = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    p = map_dbl(t_res, ~ .x$p.value),
    stars = p_to_stars(p)
  )

# Star positions for risky vs sure
star_positions <- rating_summary %>%
  group_by(test_part, option_complexity) %>%
  summarise(
    y = 6.5,
    .groups = "drop"
  ) %>%
  mutate(x = option_complexity) %>%
  left_join(t_tests_risky_sure, by = c("test_part","option_complexity"))

# ============================================================
# SIMPLE vs COMPLEX t-tests (per domain, collapsed risky/sure)
# ============================================================

t_simple_complex <- rating %>%
  group_by(test_part, subject) %>%
  summarise(
    simple  = mean(confidence_rating[option_complexity == "simple"]),
    complex = mean(confidence_rating[option_complexity == "complex"]),
    .groups = "drop"
  ) %>%
  group_by(test_part) %>%
  summarise(
    t_res = list(t.test(simple, complex, paired = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    p = map_dbl(t_res, ~ .x$p.value),
    stars = p_to_stars(p)
  )
```

```{r certainty-rating-tables}
# ============================================================
# Risky vs Sure t-tests (simple & complex)
# ============================================================

t_risky_sure_table <- rating %>%
  group_by(test_part, option_complexity, subject) %>%
  summarise(
    risky  = mean(confidence_rating[option_risky_level == "Risky (<100%)"]),
    sure   = mean(confidence_rating[option_risky_level == "Sure (100%)"]),
    .groups="drop"
  ) %>%
  group_by(test_part, option_complexity) %>%
  summarise(
    t_out = list(t.test(risky, sure, paired = TRUE)),
    mean_diff = mean(risky - sure),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(t_out, tidy)
  ) %>%
  unnest(tidy_res) %>%
  mutate(
    comparison = paste0("Risky vs Sure (", option_complexity, ")"),
    p_stars = p_to_stars(p.value)
  ) %>%
  select(
    test_part,
    comparison,
    mean_diff,
    t = statistic,
    df = parameter,
    p = p.value,
    p_stars
  )

# ============================================================
# Simple vs Complex t-tests (collapsed over risky/sure)
# ============================================================

t_simple_complex_table <- rating %>%
  group_by(test_part, subject) %>%
  summarise(
    simple  = mean(confidence_rating[option_complexity == "simple"]),
    complex = mean(confidence_rating[option_complexity == "complex"]),
    .groups="drop"
  ) %>%
  group_by(test_part) %>%
  summarise(
    t_out = list(t.test(simple, complex, paired = TRUE)),
    mean_diff = mean(simple - complex),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(t_out, tidy)
  ) %>%
  unnest(tidy_res) %>%
  mutate(
    comparison = "Simple vs Complex",
    p_stars = p_to_stars(p.value)
  ) %>%
  select(
    test_part,
    comparison,
    mean_diff,
    t = statistic,
    df = parameter,
    p = p.value,
    p_stars
  )

# ============================================================
# Combine all tests in one table
# ============================================================

t_summary_table <- bind_rows(
  t_risky_sure_table,
  t_simple_complex_table
) %>%
  arrange(test_part, comparison)  %>%
  mutate(table_label = "Certainty rating t-test summary")

t_summary_table %>%
  kable(digits = 4, caption = "Certainty rating t-tests")
```

```{r certainty-rating-plot, fig.width=7, fig.height=5}
# ============================================================
# 1. Define Bracket Positions for Simple vs Complex
# ============================================================

# Define the height for the brackets (slightly above the data)
bracket_y <- 7
tick_height <- 0.2

# Star labels and positions (centered between 1 and 2)
star_positions_sc <- rating_summary %>%
  group_by(test_part) %>%
  summarise(
    y = bracket_y + 0.3, # Star sits slightly above the line
    x = 1.5,             # Midpoint between Simple (1) and Complex (2)
    .groups="drop"
  ) %>%
  left_join(t_simple_complex, by = "test_part")

# Define the line segments for the brackets
# We need a horizontal line from 1 to 2, and vertical ticks at 1 and 2
brackets_sc <- rating_summary %>%
  distinct(test_part) %>%
  mutate(
    x_start = 1,
    x_end = 2,
    y = bracket_y,
    y_tick_start = bracket_y - tick_height
  )

# ============================================================
# 2. Final plot
# ============================================================

plot_certainty <- ggplot() +

  # jitter grouped inside dodge
  geom_point(
    data = rating,
    aes(
      x = option_complexity,
      y = confidence_rating,
      color = option_risky_level
    ),
    position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.15),
    alpha = 0.35,
    size = 1.7
  ) +

  # bars
  geom_col(
    data = rating_summary,
    aes(
      x = option_complexity,
      y = mean_certainty,
      fill = option_risky_level
    ),
    position = position_dodge(width = 0.6),
    width = 0.55,
    color = "black",
    alpha = 0.45
  ) +

  # CI
  geom_errorbar(
    data = rating_summary,
    aes(
      x = option_complexity,
      ymin = lower, ymax = upper,
      color = option_risky_level
    ),
    width = 0.12,
    linewidth = 0.9,
    position = position_dodge(width = 0.6)
  ) +

  # ------------------------------------------------------------
  # Significance brackets (Simple vs Complex)
  # ------------------------------------------------------------

  # Horizontal line
  geom_segment(
    data = brackets_sc,
    aes(x = x_start, xend = x_end, y = y, yend = y),
    color = "black", linewidth = 0.8
  ) +

  # Left vertical tick
  geom_segment(
    data = brackets_sc,
    aes(x = x_start, xend = x_start, y = y_tick_start, yend = y+0.03),
    color = "black", linewidth = 0.8
  ) +

  # Right vertical tick
  geom_segment(
    data = brackets_sc,
    aes(x = x_end, xend = x_end, y = y_tick_start, yend = y+0.03),
    color = "black", linewidth = 0.8
  ) +

  # Stars: Simple vs Complex (Now sitting on the bracket)
  geom_text(
    data = star_positions_sc,
    aes(x = x, y = y, label = stars),
    size = 6,
    fontface = "bold"
  ) +

  # ------------------------------------------------------------

  # Stars: Risky vs Sure (Comparison within complexity level)
  geom_text(
    data = star_positions,
    aes(x = option_complexity, y = 6.2, label = stars), # Adjusted y down slightly to fit under main bracket
    size = 5,
    fontface = "italic"
  ) +

  facet_wrap(~ test_part) +
  labs(
    x = "Option complexity",
    y = "Mean certainty rating",
    fill = "Lottery type",
    color = "Lottery type",
    title = "Cognitive uncertainty ratings across conditions"
  ) +
  # Expanded Y limit to make room for the bracket
  coord_cartesian(ylim = c(1, 7.5)) +
  theme_pub_top()

plot_certainty

ggsave("s2_plot_certainty_rating_risky_safe.png",
       plot_certainty, width = 7, height = 5, dpi = 600)
```

## Perceived complexity ratings

```{r complexity-rating-data}
# Perceived-complexity ratings, analysed analogously to the certainty
# ratings above. Ratings are rescaled +1 to the 1-7 scale (a constant
# shift that does not affect any difference score or test statistic).

rating_cplx <- data_filtered %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid) %>%
  filter(option_complexity %in% c("complex", "simple")) %>%
  mutate(
    complexity_rating = complexity_rating + 1,
    option_risky_level = case_when(
      option_certainty == "sure" ~ "Sure (100%)",
      option_certainty == "uncertain" ~ "Risky (<100%)",
      TRUE ~ option_certainty
    ),
    option_risky_level = factor(option_risky_level,
                                levels = c("Risky (<100%)","Sure (100%)")),
    option_complexity  = factor(option_complexity,
                                levels = c("simple","complex"))
  )

# Group means + 95% CI per domain x complexity x risky/sure
cplx_summary <- rating_cplx %>%
  group_by(test_part, option_complexity, option_risky_level) %>%
  summarise(
    mean_cplx = mean(complexity_rating),
    se = sd(complexity_rating)/sqrt(n()),
    lower = mean_cplx - 1.96*se,
    upper = mean_cplx + 1.96*se,
    .groups = "drop"
  )

# Risky vs Sure paired t-tests (per domain x complexity)
t_risky_sure_cplx <- rating_cplx %>%
  group_by(test_part, option_complexity, subject) %>%
  summarise(
    risky  = mean(complexity_rating[option_risky_level == "Risky (<100%)"]),
    sure   = mean(complexity_rating[option_risky_level == "Sure (100%)"]),
    .groups="drop"
  ) %>%
  group_by(test_part, option_complexity) %>%
  summarise(
    t_out = list(t.test(risky, sure, paired = TRUE)),
    mean_diff = mean(risky - sure),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(t_out, tidy)
  ) %>%
  unnest(tidy_res)

# Simple vs Complex paired t-tests (per domain, collapsed over risky/sure)
t_simple_complex_cplx <- rating_cplx %>%
  group_by(test_part, subject) %>%
  summarise(
    simple  = mean(complexity_rating[option_complexity == "simple"]),
    complex = mean(complexity_rating[option_complexity == "complex"]),
    .groups="drop"
  ) %>%
  group_by(test_part) %>%
  summarise(
    t_out = list(t.test(simple, complex, paired = TRUE)),
    mean_diff = mean(simple - complex),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(t_out, tidy)
  ) %>%
  unnest(tidy_res) %>%
  mutate(stars = p_to_stars(p.value))

# Star positions for the risky-vs-sure comparison in the plot
star_positions_cplx <- cplx_summary %>%
  distinct(test_part, option_complexity) %>%
  left_join(
    t_risky_sure_cplx %>%
      mutate(stars = p_to_stars(p.value)) %>%
      select(test_part, option_complexity, stars),
    by = c("test_part", "option_complexity")
  ) %>%
  mutate(y = 6.0)
```

```{r complexity-rating-tables}
# TABLE: Risky vs Sure + Simple vs Complex

t_risky_sure_table_cplx <- t_risky_sure_cplx %>%
  mutate(
    comparison = paste0("Risky vs Sure (", option_complexity, ")"),
    p_stars = p_to_stars(p.value)
  ) %>%
  select(test_part, comparison, mean_diff, statistic, parameter, p.value, p_stars)

t_simple_complex_table_cplx <- t_simple_complex_cplx %>%
  mutate(
    comparison = "Simple vs Complex",
    p_stars = p_to_stars(p.value)
  ) %>%
  select(test_part, comparison, mean_diff, statistic, parameter, p.value, p_stars)

complexity_t_summary_table <- bind_rows(
  t_risky_sure_table_cplx,
  t_simple_complex_table_cplx
) %>%
  arrange(test_part, comparison) %>%
  mutate(table_label = "Complexity rating t-test summary")

complexity_t_summary_table %>%
  kable(digits = 4, caption = "Complexity rating t-tests")
```

```{r complexity-rating-plot, fig.width=7, fig.height=5}
# ============================================================
# 1. Define Bracket Positions for Simple vs Complex
# ============================================================

# Define height for the brackets (needs to be high enough to clear the 'Complex' bars)
bracket_y_cplx <- 6.5
tick_height <- 0.2

# Update star positions to sit on the bracket
star_positions_sc_cplx <- cplx_summary %>%
  group_by(test_part) %>%
  summarise(
    y = bracket_y_cplx + 0.3,
    x = 1.5,
    .groups="drop"
  ) %>%
  left_join(t_simple_complex_cplx, by = "test_part")

# Define the line segments for the brackets
brackets_sc_cplx <- cplx_summary %>%
  distinct(test_part) %>%
  mutate(
    x_start = 1,
    x_end = 2,
    y = bracket_y_cplx,
    y_tick_start = bracket_y_cplx - tick_height
  )

# ============================================================
# 2. Final Plot
# ============================================================

plot_complexity <- ggplot() +

  # Data points (jittered)
  geom_point(
    data = rating_cplx,
    aes(x = option_complexity, y = complexity_rating,
        color = option_risky_level),
    position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.15),
    alpha = 0.35,
    size = 1.7
  ) +

  # Bars
  geom_col(
    data = cplx_summary,
    aes(x = option_complexity, y = mean_cplx, fill = option_risky_level),
    position = position_dodge(width = 0.6),
    width = 0.55,
    color = "black",
    alpha = 0.45
  ) +

  # Error bars
  geom_errorbar(
    data = cplx_summary,
    aes(x = option_complexity, ymin = lower, ymax = upper,
        color = option_risky_level),
    position = position_dodge(width = 0.6),
    width = 0.12,
    linewidth = 0.9
  ) +

  # ------------------------------------------------------------
  # Significance brackets (Simple vs Complex)
  # ------------------------------------------------------------

  # Horizontal line
  geom_segment(
    data = brackets_sc_cplx,
    aes(x = x_start, xend = x_end, y = y, yend = y),
    color = "black", linewidth = 0.8
  ) +

  # Left vertical tick
  geom_segment(
    data = brackets_sc_cplx,
    aes(x = x_start, xend = x_start, y = y_tick_start, yend = y+0.03),
    color = "black", linewidth = 0.8
  ) +

  # Right vertical tick
  geom_segment(
    data = brackets_sc_cplx,
    aes(x = x_end, xend = x_end, y = y_tick_start, yend = y+0.03),
    color = "black", linewidth = 0.8
  ) +

  # Stars: Simple vs Complex (Now sitting on the bracket)
  geom_text(
    data = star_positions_sc_cplx,
    aes(x = x, y = y, label = stars),
    size = 6,
    fontface = "bold"
  ) +

  # ------------------------------------------------------------

  # Stars: Risky vs Sure (Comparison within complexity level)
  # Ensure these don't overlap with the bracket; adjusted y position if needed
  geom_text(
    data = star_positions_cplx,
    aes(x = option_complexity, y = y, label = stars),
    size = 5,
    fontface = "italic"
  ) +

  facet_wrap(~ test_part) +
  labs(
    x = "Option complexity",
    y = "Mean complexity rating",
    fill = "Lottery type",
    color = "Lottery type",
    title = "Complexity ratings across conditions"
  ) +
  # Expanded Y limit to accommodate the higher bracket
  coord_cartesian(ylim = c(1, 7)) +
  theme_pub_top()

plot_complexity

ggsave("s2_plot_complexity_rating_risky_safe.png",
       plot_complexity, width = 7, height = 5, dpi = 600)
```

## Rating differences and complex-option choice

Spearman correlations assess (a) the association between rating difference
scores (complex - simple) for perceived complexity and the proportion of
complex choices, and (b) the association between rating difference scores
for cognitive uncertainty (complex - simple) and the proportion of complex
choices, separately for gains and losses. The gain-domain panel is reported
in the main text; the loss-domain panel is an Appendix A output.

```{r rating-diff-scores}
rating <- data_filtered %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid) %>%
  filter(option_complexity %in% c("complex", "simple")) %>%
  mutate(certainty_rating = confidence_rating)

rating_complex_summary <- rating %>%
  group_by(prolific_pid, option_complexity) %>%
  summarise(
    mean_complex_rating = mean(complexity_rating, na.rm = TRUE),
    .groups = "drop"
  )

rating_certainty_summary <- rating %>%
  group_by(prolific_pid, option_complexity) %>%
  summarise(
    mean_certainty_rating = mean(certainty_rating, na.rm = TRUE),
    .groups = "drop"
  )

# --- Prepare diff scores ---
rating_complex_diff <- rating_complex_summary %>%
  pivot_wider(
    id_cols = prolific_pid,
    names_from = option_complexity,
    values_from = mean_complex_rating
  ) %>%
  mutate(diff_complexity = complex - simple) %>%
  select(prolific_pid, diff_complexity)

# Cognitive uncertainty = reversed certainty, hence the sign flip
rating_certainty_diff <- rating_certainty_summary %>%
  pivot_wider(
    id_cols = prolific_pid,
    names_from = option_complexity,
    values_from = mean_certainty_rating
  ) %>%
  mutate(diff_certainty = (complex - simple) * -1) %>%
  select(prolific_pid, diff_certainty)

# --- Join with complex-choice proportions ---
joined_corr_df <- complex_choice_summary_individual %>%
  left_join(rating_complex_diff,  by = "prolific_pid") %>%
  left_join(rating_certainty_diff, by = "prolific_pid") %>%
  mutate(test_part = factor(test_part, levels = c("Gain", "Loss")))
```

```{r rating-diff-choice-correlations}
# Correlation summary table (the same Spearman tests that are annotated in
# the panels below)
corr_rating_choice <- joined_corr_df %>%
  group_by(test_part) %>%
  summarise(
    `Perceived complexity` = list(tidy(
      cor.test(diff_complexity, prop_complex, method = "spearman"))),
    `Cognitive uncertainty` = list(tidy(
      cor.test(diff_certainty, prop_complex, method = "spearman"))),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(`Perceived complexity`, `Cognitive uncertainty`),
    names_to = "predictor",
    values_to = "res"
  ) %>%
  unnest(res) %>%
  mutate(stars = p_to_stars(p.value)) %>%
  select(test_part, predictor, estimate, statistic, p.value, stars)

corr_rating_choice %>%
  kable(digits = 4,
        caption = "Spearman correlations: rating difference (complex − simple) vs complex-choice proportion (Gain = main text; Loss = Appendix A)")
```

```{r rating-diff-choice-plots, fig.width=7, fig.height=5}
# =========================================================
# FUNCTION: build one plot for a given domain (Gain or Loss)
# =========================================================
make_corr_plot_for_domain <- function(domain_label, df) {

  df_domain <- df %>% filter(test_part == domain_label)

  # 1) correlations for this domain
  corr_complexity <- df_domain %>%
    summarise(res = list(cor.test(diff_complexity, prop_complex, method = "spearman"))) %>%
    mutate(tidy = map(res, tidy)) %>%
    unnest(tidy) %>%
    mutate(predictor = "Perceived complexity")

  corr_certainty <- df_domain %>%
    summarise(res = list(cor.test(diff_certainty, prop_complex, method = "spearman"))) %>%
    mutate(tidy = map(res, tidy)) %>%
    unnest(tidy) %>%
    mutate(predictor = "Cognitive uncertainty")

  corr_all <- bind_rows(corr_complexity, corr_certainty) %>%
    select(predictor, estimate, p.value) %>%
    mutate(
      stars = p_to_stars(p.value),
      label = paste0("ρ = ", round(estimate, 2), " ", stars)
    )

  # 2) long df for plotting
  plot_df <- df_domain %>%
    select(prolific_pid, prop_complex, diff_complexity, diff_certainty) %>%
    pivot_longer(
      cols = c(diff_complexity, diff_certainty),
      names_to = "predictor",
      values_to = "diff_score"
    ) %>%
    mutate(
      predictor = recode(
        predictor,
        diff_complexity = "Perceived complexity",
        diff_certainty  = "Cognitive uncertainty"
      )
    )

  plot_df$predictor <- factor(
    plot_df$predictor,
    levels = c("Perceived complexity", "Cognitive uncertainty")
  )
  corr_all$predictor <- factor(corr_all$predictor, levels = levels(plot_df$predictor))

  # 3) plot (2×1 panel)
  ggplot(plot_df, aes(x = diff_score, y = prop_complex)) +
    geom_point(alpha = 0.55) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 1) +
    facet_grid(predictor ~ .) +
    geom_text(
      data = corr_all,
      aes(x = -Inf, y = Inf, label = label),
      hjust = -0.05, vjust = 1.1, size = 7, fontface = "bold"
    ) +
    labs(
      x = "Rating difference (complex – simple)",
      y = "Proportion choosing complex option",
      title = paste0("Correlations between rating diff and complex choices (", domain_label, ")")
    ) +
    theme_pub()
}

# =========================================================
# MAKE TWO SEPARATE PLOTS (Gain = main text; Loss = Appendix A)
# =========================================================
plot_gain <- make_corr_plot_for_domain("Gain", joined_corr_df)
plot_loss <- make_corr_plot_for_domain("Loss", joined_corr_df)

plot_gain
plot_loss

# =========================================================
# SAVE
# =========================================================
ggsave(
  filename = "s2_plot_corr_rating_diff_complex_gain.png",
  plot = plot_gain,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)

ggsave(
  filename = "s2_plot_corr_rating_diff_complex_loss.png",
  plot = plot_loss,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)
```

# Cognitive-ability correlations

Cognitive ability (CA) is the mean accuracy across the Berlin Numeracy Test
(BNT) and the Health & Medical Test (HMT). Spearman correlations relate CA
to complex-option choice and to EV-consistent choice, per domain. The
loss-domain CA plot is an Appendix A output.

```{r cognitive-ability-scores}
# Cognitive-ability score per participant: mean accuracy across the two
# tasks (BNT and HMT)
result_ca <- data %>%
  group_by(prolific_pid) %>%
  dplyr::summarize(
    accuracy_BNT = mean(accuracy_BNT, na.rm = TRUE),
    accuracy_HMT = mean(accuracy_HMT, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(accuracy_BNT, accuracy_HMT), names_to = "accuracy_type", values_to = "score")

result_ca <- result_ca %>%
  group_by(prolific_pid) %>%
  dplyr::summarize(ca_results = mean(score, na.rm = TRUE))
```

## CA and complex-option choice

```{r ca-complex-choice, fig.width=7, fig.height=5}
# --- join data ---
joined_df <- inner_join(result_ca, complex_choice_summary_individual,
                        by = "prolific_pid") %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  mutate(test_part = factor(test_part, levels = c("Gain", "Loss")))

# =========================================================
# FUNCTION: build one plot for a given domain (Gain or Loss)
# =========================================================
make_ca_corr_plot_for_domain <- function(domain_label, df) {

  df_domain <- df %>% filter(test_part == domain_label)

  # Spearman correlation (for annotation + table output)
  corr_res <- tidy(cor.test(df_domain$ca_results, df_domain$prop_complex, method = "spearman")) %>%
    mutate(
      test_part = domain_label,
      table_label = "CA correlation with complex choice",
      stars = p_to_stars(p.value),
      label = paste0("ρ = ", sprintf("%.2f", estimate), "  ", stars)
    )

  # plot
  p <- ggplot(df_domain, aes(x = ca_results, y = prop_complex)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
    geom_text(
      data = corr_res,
      aes(x = -Inf, y = Inf, label = label),
      hjust = -0.05, vjust = 1.1,
      size = 7, fontface = "bold",
      inherit.aes = FALSE
    ) +
    labs(
      x = "CA performance",
      y = "Proportion of complex choices",
      title = paste0("Correlation between CA and complex choice (", domain_label, ")")
    ) +
    theme_pub(base_size = 14)

  list(plot = p, corr = corr_res)
}

# =========================================================
# MAKE TWO SEPARATE PLOTS + RESULTS
# =========================================================
out_gain <- make_ca_corr_plot_for_domain("Gain", joined_df)
out_loss <- make_ca_corr_plot_for_domain("Loss", joined_df)

plot_ca_corr_gain <- out_gain$plot
plot_ca_corr_loss <- out_loss$plot

corr_results_gain <- out_gain$corr
corr_results_loss <- out_loss$corr
corr_results <- bind_rows(corr_results_gain, corr_results_loss)

corr_results %>%
  select(test_part, table_label, estimate, statistic, p.value, stars) %>%
  kable(digits = 4,
        caption = "Spearman correlations: CA vs complex-choice proportion (Loss = Appendix A)")

plot_ca_corr_gain
plot_ca_corr_loss

# =========================================================
# SAVE (gain = manuscript figure panel; loss = Appendix A)
# =========================================================
ggsave(
  filename = "s2_plot_ca_corr_gain.png",
  plot = plot_ca_corr_gain,
  width = 7,
  height = 5,
  units = "in",
  dpi = 600
)

ggsave(
  filename = "s2_plot_ca_corr_loss.png",
  plot = plot_ca_corr_loss,
  width = 7,
  height = 5.0,
  units = "in",
  dpi = 600
)
```

## CA and EV-consistent choice

```{r ca-ev-consistency}
# --- 1. Compute EV-consistent choice per trial ---
df_ev <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  filter(EV_diff_level != 0) %>%              # ignore EV=0 trials
  mutate(
    domain = factor(ifelse(test_part == "Gain", "Gains", "Losses"),
                    levels = c("Gains", "Losses")),

    # which option has higher EV?
    higher_EV_option = ifelse(EV_diff_level > 0, "A", "B"),

    # chosen option (f = left/A, j = right/B)
    chosen_option = ifelse(real_response == "f", "A", "B"),

    # EV-consistent = chose higher EV option
    ev_consistent = ifelse(chosen_option == higher_EV_option, 1, 0)
  )


# --- 2. Proportion per participant × domain ---
ev_consistency_individual <- df_ev %>%
  group_by(prolific_pid, domain) %>%
  summarise(
    prop_ev_consistent = mean(ev_consistent, na.rm = TRUE),
    .groups = "drop"
  )


# --- 3. Join with CA scores ---
joined_ev <- inner_join(result_ca, ev_consistency_individual, by = "prolific_pid")


# --- 4. Spearman correlations separately for Gains and Losses ---
corr_ev_results <- joined_ev %>%
  group_by(domain) %>%
  summarise(
    cor_test = list(tidy(cor.test(ca_results,
                                  prop_ev_consistent,
                                  method = "spearman"))),
    .groups = "drop"
  ) %>%
  unnest(cor_test) %>%
  mutate(table_label = "CA correlation with EV-consistent choice")

corr_ev_results %>%
  kable(digits = 4,
        caption = "Spearman correlations: CA vs EV-consistent choice proportion")

# Descriptive check (mean EV-consistency per domain); not a reported
# statistic in the paper — retained for reference
ev_consistency_individual %>%
  group_by(domain) %>%
  summarise(mean_prop_ev_consistent = mean(prop_ev_consistent)) %>%
  kable(digits = 4, caption = "Mean EV-consistent choice proportion by domain")
```

```{r ca-ev-consistency-plot, fig.width=7, fig.height=5}
# add stars and clean label formatting
label_df_ev <- corr_ev_results %>%
  mutate(
    stars = p_to_stars(p.value),
    label = paste0("ρ = ",
                   sprintf("%.2f", estimate),
                   "  ", stars)
  )

# --- APA-style plot with stars ---
plot_ev_corr <- ggplot(joined_ev,
                       aes(x = ca_results, y = prop_ev_consistent)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = T, linewidth = 0.8) +
  facet_wrap(~ domain) +
  geom_text(data = label_df_ev,
            aes(label = label),
            x = -Inf, y = Inf,
            hjust = -0.05, vjust = 1.1,
            size = 7,
            fontface = "bold") +
  labs(
    x = "CA performance",
    y = "Proportion of EV-consistent choices",
    title = "Correlation between CA and EV-consistent choice proportion"
  ) +
  theme_pub(base_size = 14)

plot_ev_corr

ggsave(
  filename = "s2_plot_ev_corr.png",
  plot = plot_ev_corr,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# High-EV-difference trial accuracy

Catch trials had a large EV difference (+/- 70); accuracy is the proportion
choosing the EV-better option (choice rule reversed in the loss domain).

```{r catch-accuracy-data}
# Catch trials: large EV difference (+/- 70). Accuracy = choosing the
# EV-better option (choice rule reversed in the loss domain)
df_test2 <- cleaned_RT_data %>%
  filter(trial_type_label == "test" & skew == "catch")

df_test2 <- df_test2 %>%
  mutate(
    real_response = case_when(
      risk_index == 1  & response == "f" ~ "f",
      risk_index == 1  & response == "j" ~ "j",
      risk_index == -1 & response == "f" ~ "j",
      risk_index == -1 & response == "j" ~ "f"
    ),
    good_decision = case_when(
      # --- Gain_Catch: same rule as before ---
      test_part == "Gain_Catch" & EV_diff_level == 70  & real_response == "f" ~ 1,
      test_part == "Gain_Catch" & EV_diff_level == -70 & real_response == "j" ~ 1,

      # --- Loss_Catch: reverse rule ---
      test_part == "Loss_Catch" & EV_diff_level == 70  & real_response == "j" ~ 1,
      test_part == "Loss_Catch" & EV_diff_level == -70 & real_response == "f" ~ 1,

      TRUE ~ 0
    )
  )

catch_accuracy_individual <- df_test2 %>%
  group_by(prolific_pid, test_part) %>%
  summarise(
    n_trials = n(),
    prop_catch_accuracy = mean(good_decision),
    .groups = "drop"
  )

catch_accuracy_group <-  catch_accuracy_individual %>%
  group_by(test_part) %>%
  summarise(
    mean_catch_prop = mean(prop_catch_accuracy, na.rm = TRUE),
    se = sd(prop_catch_accuracy, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_catch_prop - 1.96 * se,
    upper = mean_catch_prop + 1.96 * se
  )

catch_accuracy_group %>%
  kable(digits = 4, caption = "High-EV-difference trial accuracy by domain")
```

```{r catch-accuracy-tests}
# ---------- One-sample (vs .5) and paired t-tests ----------
t_catch_gain <- t.test(
  catch_accuracy_individual$prop_catch_accuracy[
    catch_accuracy_individual$test_part == "Gain_Catch"
  ],
  mu = 0.5, alternative = "two.sided"
)

t_catch_loss <- t.test(
  catch_accuracy_individual$prop_catch_accuracy[
    catch_accuracy_individual$test_part == "Loss_Catch"
  ],
  mu = 0.5, alternative = "two.sided"
)

# Paired: same participants in Gain & Loss
catch_wide <- catch_accuracy_individual %>%
  pivot_wider(
    id_cols    = prolific_pid,
    names_from = test_part,
    values_from = prop_catch_accuracy
  )

t_catch_gain_vs_loss <- t.test(
  catch_wide$Gain_Catch, catch_wide$Loss_Catch,
  paired = TRUE, alternative = "two.sided"
)

# Summary table
t_summary_catch <- bind_rows(
  tidy(t_catch_gain)         %>% mutate(test = "Gain vs 0.5"),
  tidy(t_catch_loss)         %>% mutate(test = "Loss vs 0.5"),
  tidy(t_catch_gain_vs_loss) %>% mutate(test = "Gain vs Loss (paired)")
) %>%
  select(test, estimate, statistic, parameter, p.value, conf.low, conf.high)

t_summary_catch %>%
  kable(digits = 4, caption = "High-EV-difference trial accuracy t-tests")
```

```{r catch-accuracy-plot, fig.width=7, fig.height=5}
stars_gain <- p_to_stars(t_catch_gain$p.value)
stars_loss <- p_to_stars(t_catch_loss$p.value)
stars_pair <- p_to_stars(t_catch_gain_vs_loss$p.value)

# ---------- Prepare data for plotting ----------
plot_pts <- catch_accuracy_individual %>%
  mutate(
    test_part = factor(
      test_part,
      levels = c("Gain_Catch", "Loss_Catch"),
      labels = c("Gains", "Losses")
    )
  )

plot_bars <- catch_accuracy_group %>%
  mutate(
    test_part = factor(
      test_part,
      levels = c("Gain_Catch", "Loss_Catch"),
      labels = c("Gains", "Losses")
    )
  )

# fixed vertical positions (all stars above the dots)
y_gain_star <- 1.02
y_loss_star <- 1.02
bracket_y   <- 1.04
label_y     <- 1.045

# ---------- Plot ----------
plot_catch_accuracy <- ggplot() +
  # individual participants
  geom_jitter(
    data = plot_pts,
    aes(x = test_part, y = prop_catch_accuracy),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  # group mean bars (semi-transparent)
  geom_col(
    data = plot_bars,
    aes(x = test_part, y = mean_catch_prop),
    width = 0.55,
    fill  = "grey60",
    color = "black",
    alpha = 0.45
  ) +
  # 95% CI (bold)
  geom_errorbar(
    data = plot_bars,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.12,
    linewidth = 0.9
  ) +
  # chance line at 0.5
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  # one-sample stars above each bar
  geom_text(
    aes(x = "Gains",  y = y_gain_star, label = stars_gain),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  geom_text(
    aes(x = "Losses", y = y_loss_star, label = stars_loss),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  # bracket + star for paired Gain vs Loss
  geom_segment(aes(x = 1, xend = 1, y = bracket_y - 0.01, yend = bracket_y)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y - 0.01, yend = bracket_y)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y,       yend = bracket_y)) +
  geom_text(
    aes(x = 1.5, y = label_y, label = stars_pair),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  scale_y_continuous(
    limits = c(0, 1.1),
    breaks = seq(0, 1, by = 0.25),
    expand = c(0, 0)
  ) +
  labs(
    x = "Domain",
    y = "High EV diff trial accuracy (proportion correct)",
    title = "High EV diff trial accuracy in gain and loss domains"
  ) +
  theme_pub()

plot_catch_accuracy

ggsave(
  filename = "s2_plot_catch_accuracy.png",
  plot = plot_catch_accuracy,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# Appendix B: preregistered complex-role test

Proportion of complex-option choices when the risky vs the sure option is
the complex one, per domain (figure fig:Preregistered in the paper).

```{r complex-role-data}
complex_choice_summary_individual2 <- df_test1 %>%
  mutate(complex_role = ifelse(complexity_option == "A", "Risky option complex", "Sure option complex")) %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by(prolific_pid, test_part, complex_role) %>%
  summarise(
    n_trials = n(),
    n_complex = sum(complex_choice, na.rm = TRUE),
    prop_complex = mean(complex_choice, na.rm = TRUE),
    .groups = "drop"
  )

complex_choice_summary_group2 <-  complex_choice_summary_individual2 %>%
  group_by(test_part, complex_role) %>%
  summarise(
    mean_complex_prop = mean(prop_complex, na.rm = TRUE),
    se = sd(prop_complex, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_complex_prop - 1.96 * se,
    upper = mean_complex_prop + 1.96 * se
  )
```

```{r complex-role-plot-data}
# -------------------------------------------------------------------
# 1) Prepare plotting data (facet by domain, x = which option is complex)
# -------------------------------------------------------------------
plot_df_points <- complex_choice_summary_individual2 %>%
  mutate(
    test_part = factor(test_part, levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses")),
    complex_role = factor(
      complex_role,
      levels = c("Risky option complex", "Sure option complex")
    )
  )

plot_df_bars <- complex_choice_summary_group2 %>%
  mutate(
    test_part = factor(test_part, levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses")),
    complex_role = factor(
      complex_role,
      levels = c("Risky option complex", "Sure option complex")
    )
  )

# -------------------------------------------------------------------
# 2) One-sample t-tests vs 0.5 (for each domain × role)
# -------------------------------------------------------------------
tests_one <- plot_df_points %>%
  group_by(test_part, complex_role) %>%
  summarise(
    p = t.test(prop_complex, mu = 0.5)$p.value,
    .groups = "drop"
  ) %>%
  mutate(
    stars = p_to_stars(p),
    x = as.numeric(complex_role),
    y = 1.01  # position of stars vs 0.5
  )

# -------------------------------------------------------------------
# 3) Paired t-tests within domain: risky complex vs sure complex
# -------------------------------------------------------------------
paired_tests <- complex_choice_summary_individual2 %>%
  select(prolific_pid, test_part, complex_role, prop_complex) %>%
  pivot_wider(
    names_from  = complex_role,
    values_from = prop_complex
  ) %>%
  mutate(
    test_part = factor(test_part, levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses"))
  ) %>%
  group_by(test_part) %>%
  summarise(
    p = t.test(`Risky option complex`,
               `Sure option complex`,
               paired = TRUE)$p.value,
    .groups = "drop"
  ) %>%
  mutate(
    stars   = p_to_stars(p),
    x_start = 1,
    x_end   = 2,
    y       = 1.07  # bracket height
  )
```

```{r complex-role-tables}
# ============================================================
# 1) One-sample t-tests vs 0.5
# ============================================================

one_sample_results <- plot_df_points %>%
  group_by(test_part, complex_role) %>%
  summarise(
    t_res = list(t.test(prop_complex, mu = 0.5)),
    mean_prop = mean(prop_complex),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(t_res, tidy),
    t = map_dbl(tidy_res, "statistic"),
    df = map_dbl(tidy_res, "parameter"),
    p = map_dbl(tidy_res, "p.value"),
    p_stars = p_to_stars(p),
    comparison = paste0(complex_role, " vs 0.5")
  ) %>%
  select(test_part, comparison, mean_prop, t, df, p, p_stars)

# ============================================================
# 2) Paired test: Risky option complex vs Sure option complex
# ============================================================

paired_results <- complex_choice_summary_individual2 %>%
  select(prolific_pid, test_part, complex_role, prop_complex) %>%
  pivot_wider(
    names_from = complex_role,
    values_from = prop_complex
  ) %>%
  mutate(
    test_part = factor(test_part,
                       levels = c("Gain","Loss"),
                       labels = c("Gains","Losses"))
  ) %>%
  group_by(test_part) %>%
  summarise(
    t_res = list(t.test(`Risky option complex`,
                        `Sure option complex`,
                        paired = TRUE)),
    mean_diff = mean(`Risky option complex` - `Sure option complex`),
    mean_risky = mean(`Risky option complex`),
    mean_sure  = mean(`Sure option complex`),
    .groups = "drop"
  ) %>%
  mutate(
    tidy_res = map(t_res, tidy),
    t = map_dbl(tidy_res, "statistic"),
    df = map_dbl(tidy_res, "parameter"),
    p = map_dbl(tidy_res, "p.value"),
    p_stars = p_to_stars(p),
    comparison = "Risky complex vs Sure complex"
  ) %>%
  select(test_part, comparison,
         mean_risky, mean_sure, mean_diff,
         t, df, p, p_stars)

# ============================================================
# 3) Combine into one summary table
# ============================================================

complex_choice_test_summary <- bind_rows(
  one_sample_results,
  paired_results
) %>%
  arrange(test_part, comparison)

complex_choice_test_summary %>%
  kable(digits = 4,
        caption = "Complex-option choice by complex-option role (Appendix B)")
```

```{r complex-role-plot, fig.width=7, fig.height=5}
pair_label_df <- paired_tests %>%
  mutate(
    x = 1.5,
    y = 1.05  # label height
  )

# -------------------------------------------------------------------
# 4) Plot (faceted by domain)
# -------------------------------------------------------------------
plot_complex_role <- ggplot() +
  # individual participants
  geom_jitter(
    data = plot_df_points,
    aes(x = complex_role, y = prop_complex),
    width = 0.10, height = 0,
    alpha = 0.45, size = 1.8
  ) +
  # mean bars (semi-transparent so dots remain visible)
  geom_col(
    data = plot_df_bars,
    aes(x = complex_role, y = mean_complex_prop, fill = complex_role),
    width = 0.60,
    alpha = 0.45,
    colour = "black"
  ) +
  # 95% CI
  geom_errorbar(
    data = plot_df_bars,
    aes(x = complex_role, ymin = lower, ymax = upper, , color = "#b00000"),
    width = 0.18, linewidth = 1.3
  ) +
  # reference line at 0.5
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  # stars for one-sample tests (vs 0.5)
  geom_text(
    data = tests_one,
    aes(x = x, y = y, label = stars),
    vjust = 0, size = 6, fontface = "bold"
  ) +

  geom_segment(
    data = paired_tests,
    aes(x = x_start, xend = x_start, y = y - 0.01, yend = y)
  ) +
  geom_segment(
    data = paired_tests,
    aes(x = x_end, xend = x_end, y = y - 0.01, yend = y)
  ) +
  geom_segment(
    data = paired_tests,
    aes(x = x_start, xend = x_end, y = y, yend = y)
  ) +
  geom_text(
    data = pair_label_df,
    aes(x = x, y = y, label = stars),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  scale_y_continuous(
    limits = c(0, 1.12),
    breaks = seq(0, 1, by = 0.25),
    expand = c(0, 0)
  ) +
  scale_fill_manual(
    values = c("Risky option complex" = "grey50",
               "Sure option complex"  = "grey80")
  ) +
  labs(
    x = "Which option is complex",
    y = "Proportion of choices for complex option",
    title = "Complex-option choices when risky vs sure option is complex",
    fill = NULL
  ) +
  facet_wrap(~ test_part, nrow = 1) +
  theme_pub() +
  theme(legend.position = "none")

plot_complex_role

ggsave(
  filename = "s2_plot_complex_role.png",
  plot = plot_complex_role,
  width = 7,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# Manuscript figures

Assemble the manuscript summary figures from the panels saved as .png by
the chunks above. Every panel referenced here is generated earlier in this
document, so the assemblies work in a single fresh knit.
(Previously, some assembly variants referenced panels that no chunk saved
— plot_ca_corr.png, plot_corr_rating_diff_complex.png,
plot_complexity_rating.png, plot_certainty_rating.png; those superseded
assemblies were removed and only the reported figures are built here.)

```{r figure-panels}
# Load saved figure panels for the two manuscript summary figures
p1  <- ggdraw() + draw_image("s2_plot_complex_choice.png", scale = 1)
p3  <- ggdraw() + draw_image("s2_plot_risky_choice.png", scale = 1)
p4  <- ggdraw() + draw_image("s2_plot_rt_median_diff.png", scale = 1)
p5  <- ggdraw() + draw_image("s2_plot_rt_risky.png", scale = 1)
p6  <- ggdraw() + draw_image("s2_plot_complexity_rating_risky_safe.png", scale = 1)
p7  <- ggdraw() + draw_image("s2_plot_certainty_rating_risky_safe.png", scale = 1)
p8  <- ggdraw() + draw_image("s2_plot_corr_rating_diff_complex_gain.png", scale = 1)
p9  <- ggdraw() + draw_image("s2_plot_catch_accuracy.png", scale = 1)
p10 <- ggdraw() + draw_image("s2_plot_ca_corr_gain.png", scale = 1)
p11 <- ggdraw() + draw_image("s2_plot_ev_corr.png", scale = 1)
```

```{r figure-summary-1, fig.width=12, fig.height=10}
summary_analysis1 <- plot_grid(
  p1, p3, p4, p5,

  labels = c("(a)", "(b)", "(c)", "(d)"),
  label_size = 18,
  label_fontface = "bold",
  ncol = 2,
  align = "hv"
)

# Save
ggsave(
  "summary_analysis_1_update.png",
  summary_analysis1,
  width = 12,
  height = 10,
  dpi = 600
)
```

```{r figure-summary-2, fig.width=14, fig.height=14}
summary_analysis2 <- plot_grid(
  p6, p7,
  p8, p9, p10, p11,
  labels = c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)"),
  label_size = 18,
  label_fontface = "bold",
  ncol = 2,
  align = "hv"
)

# Save
ggsave(
  "summary_analysis_2.png",
  summary_analysis2,
  width = 14,
  height = 14,
  dpi = 600
)
```
