---
title: "Study 1 — Behavioural Analysis"
output: html_document
---

```{r}
# ============================================================================
# study1_behavioural_analysis.Rmd — model-free + regression analyses, Study 1
#
# Aim:     Main behavioural analysis of Study 1 (complexity aversion in risky
#          choice across gain and loss domains, N = 150 Prolific participants).
#          The notebook is organised as a report mirroring the manuscript:
#            1.  Setup (packages, data, shared helpers)
#            2.  Preregistered exclusions and participant flow
#                (comprehension criterion; RT cleaning pipeline;
#                trial-count criterion; exclusion-flow table; final N = 138)
#            3.  Sample descriptives (participant flow, age/gender,
#                comprehension attempts, RT distribution and timeout rate,
#                bonus payments)
#            4.  Complex-option choice (one-sample t vs .5 per domain +
#                paired gain–loss; figure panel)
#            5.  Riskier-option choice (same trio of tests; figure panel)
#            6.  Response times (median RT complex − simple and
#                riskier − safer; figure panels)
#            7.  Bayesian logistic mixed-effects regression (brms, cached to
#                m_brm_study1.rds; model_parameters table for the paper)
#            8.  EV-dominant (catch) trial accuracy (trio of tests; panel)
#            9.  Cognitive-ability correlations (CA x complex choice per
#                domain — loss panel is Appendix A; CA x EV-consistent choice)
#            10. Manuscript summary figures (plot_grid assemblies)
#          Sections 2 and 3 run the exclusions before the descriptives so the
#          notebook knits top-to-bottom; all descriptives refer to the final
#          sample, as in the manuscript.
# Inputs:  ../data/study1/merged_data_study1.csv.gz
#          (pre-exclusion data; one row per jsPsych event)
# Outputs: Figure PNGs written to the working directory (s1_plot_*.png,
#          summary_explore_analysis.png, summary_main_analysis.png) and
#          m_brm_study1.rds (cached Bayesian model fit).
# Exclusions applied here:
#          1. Participants needing > 8 attempts to pass the four
#             comprehension questions.
#          2. Trials with RT < 1 s or > 30 s.
#          3. Trials with RT outside median ± 3 SD, computed per
#             participant × domain.
#          4. Participants with < 98 remaining test trials
#             (i.e. > 50% of trials lost to RT cleaning).
# Usage:   Knit/run from 02_behavioural/ so the relative data path resolves.
# ============================================================================
```

# 1. Setup

```{r}
library(data.table)
library(tidyverse)   # dplyr, tidyr, ggplot2, purrr, stringr, ...
library(jsonlite)
library(lubridate)
library(magrittr)
library(knitr)
library(broom)
library(cowplot)
```

```{r}
# Load the pre-exclusion dataset (fread reads .gz transparently)
data <- fread("../data/study1/merged_data_study1.csv.gz")
```

```{r}
# ---------- Shared APA-like plotting theme ----------
# base_size 13 for the main figure panels, 14 for the exploratory panels
theme_pub <- function(base_size = 13) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.title         = element_text(face = "bold"),
      axis.text          = element_text(color = "black"),
      strip.text         = element_text(face = "bold"),
      panel.border       = element_rect(color = "black", fill = NA, linewidth = 0.5),
      plot.title         = element_text(face = "bold", hjust = 0.5),
      axis.line          = element_line(color = "black"),
      legend.position    = "none"
    )
}

# ---------- Helper for significance stars (vectorised, NA-safe) ----------
p_to_stars <- function(p){
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ "ns"
  )
}
```

# 2. Preregistered exclusions and participant flow

## 2.1 Comprehension-question criterion

Each participant had to pass four comprehension questions; every failed block
adds question events. Participants needing more than 8 question events
(i.e. more than one full retry) are excluded.

```{r}
data_filtered <- data %>%
  group_by(prolific_pid) %>%
  mutate(attempts_to_pass_four_questions = sum(trial_type_label == "question", na.rm = TRUE)) %>%
  ungroup() %>%
  filter(attempts_to_pass_four_questions <= 8)

n_enrolled   <- length(unique(data$prolific_pid))
n_after_comp <- length(unique(data_filtered$prolific_pid))
```

## 2.2 RT cleaning (trial level)

```{r}
# Keep non-practice test trials, convert RT from ms to s
cleaned_RT_data <- data_filtered %>%
  filter(trial_type_label == 'test' & test_part != 'prac') %>%
  mutate(rt = as.numeric(rt_trial/1000))

# Snapshot of ALL test trials BEFORE any RT filter. The timeout rate
# (Section 3.4) must be computed from this object: after the < 1 s / > 30 s
# filter below, no trial can have rt >= 35 s, so a post-filter timeout rate
# is structurally zero.
rt_pre_filter <- cleaned_RT_data

# First-pass flag: absolute RT outliers (< 1 s or > 30 s)
cleaned_RT_data <- cleaned_RT_data %>%
  mutate(
    is_outlier1 = rt < 1 | rt > 30
  )

n_trials_start <- nrow(cleaned_RT_data)
n_abs_removed  <- sum(cleaned_RT_data$is_outlier1, na.rm = TRUE)
```

```{r}
# Second-pass outlier flag: median ± 3 SD per participant × domain
cleaned_RT_data <- cleaned_RT_data %>%
  filter(!is_outlier1 ) %>%
  group_by(prolific_pid, test_part) %>%
  mutate(
    median_rt = median(rt),
    sd_rt = sd(rt),
    is_outlier2 = rt < (median_rt - 3 * sd_rt) | rt > (median_rt + 3 * sd_rt)
  )

n_mad_removed <- sum(cleaned_RT_data$is_outlier2, na.rm = TRUE)

cleaned_RT_data <- cleaned_RT_data %>%
  filter(!is_outlier2)

n_trials_after_mad <- nrow(cleaned_RT_data)
```

## 2.3 Trial-count criterion (participant level)

```{r}
# Exclude participants with fewer than 98 (= 196/2) remaining trials,
# i.e. more than 50% of trials lost to RT cleaning
trial_counts <- cleaned_RT_data %>%
  group_by(prolific_pid) %>%
  dplyr::summarize(trials = n())


participants_to_keep <- trial_counts %>%
  filter(trials >= 196/2) %>%
  select(prolific_pid)


cleaned_RT_data <- cleaned_RT_data %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid)

n_final              <- length(unique(cleaned_RT_data$prolific_pid))
n_rt_excluded        <- n_after_comp - n_final
n_trials_dropped_ppt <- n_trials_after_mad - nrow(cleaned_RT_data)
```

## 2.4 Exclusion-flow table

```{r}
exclusion_flow <- tibble(
  Step = c(
    "Recruited",
    "Comprehension criterion (> 8 attempts)",
    "Absolute RT filter (< 1 s or > 30 s)",
    "Median ± 3 SD RT filter (per participant × domain)",
    "Trial-count criterion (< 98 trials remaining)"
  ),
  `Trials removed` = c(NA, NA, n_abs_removed, n_mad_removed, n_trials_dropped_ppt),
  `Participants removed` = c(NA, n_enrolled - n_after_comp, 0, 0, n_rt_excluded),
  `Participants remaining` = c(n_enrolled, n_after_comp, n_after_comp, n_after_comp, n_final)
)

kable(
  exclusion_flow,
  caption = "Preregistered exclusion flow: trials and participants removed at each step."
)
```

## 2.5 Analysis dataset

```{r}
# Non-catch test trials of the retained participants.
# Recode responses so "f" = risky option regardless of screen side
# (risk_index flips the mapping), then flag complex-option choices.
df_test1 <- cleaned_RT_data %>%
  filter(trial_type_label == "test" & skew != "catch") %>%
  mutate(real_response = case_when(
    risk_index == 1 & response == "f" ~ "f",
    risk_index == 1 & response == "j" ~ "j",
    risk_index == -1 & response == "f" ~ "j",
    risk_index == -1 & response == "j" ~ "f"
  )) %>%
  mutate(
    complex_choice = case_when(
      complexity_option == "A" & real_response == "f" ~ 1L,
      complexity_option == "B" & real_response == "j" ~ 1L,
      TRUE ~ 0L
    )
  )
```

# 3. Sample descriptives

## 3.1 Participant flow

```{r}
kable(
  tibble(
    Stage = c(
      "Recruited",
      "After comprehension exclusion",
      "After RT-based exclusions (final sample)"
    ),
    N = c(n_enrolled, n_after_comp, n_final)
  ),
  caption = "Participant flow."
)
```

## 3.2 Demographics

```{r}
# Parse the JSON survey responses (age, gender, education, free-text feedback)
demo_table <- data %>%
  filter(trial_type_label == 'demo')


clean_json <- function(x) {
  x <- gsub('^"|"$', "", x)             # strip outer quotes
  x <- gsub('""', '"', x, fixed = TRUE) # unescape CSV-style quotes
  x
}

demo_clean <- demo_table %>%
  mutate(
    parsed = map(response, ~{
      txt <- clean_json(.x)
      tryCatch(fromJSON(txt), error = function(e) list())
    }),
    Age       = map_chr(parsed, ~pluck(.x, "Age",       .default = NA_character_)),
    Gender    = map_chr(parsed, ~pluck(.x, "Q0",        .default = NA_character_)),
    Education = map_chr(parsed, ~pluck(.x, "Q1",        .default = NA_character_)),
    Feedback  = map_chr(parsed, ~pluck(.x, "feedback",  .default = NA_character_))
  ) %>%
  select(prolific_pid, Age, Gender, Education, Feedback)

# make Age numeric
demo_clean <- demo_clean %>%
  mutate(Age = suppressWarnings(as.numeric(Age)))

# one row per participant (if duplicated, keep first non-NA)
demo_final <- demo_clean %>%
  group_by(prolific_pid) %>%
  summarise(
    Age       = suppressWarnings(as.numeric(na.omit(Age)[1])),
    Gender    = na.omit(Gender)[1]    %||% NA_character_,
    Education = na.omit(Education)[1] %||% NA_character_,
    Feedback  = na.omit(Feedback)[1]  %||% NA_character_,
    .groups = "drop"
  )
```

```{r}
# Sample descriptives for the retained participants
demo_final_filtered <- demo_final %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid)

kable(
  demo_final_filtered %>%
    summarise(
      n = n(),
      age_mean = mean(Age, na.rm = TRUE),
      age_sd   = sd(Age, na.rm = TRUE),
      age_min  = min(Age, na.rm = TRUE),
      age_max  = max(Age, na.rm = TRUE)
    ),
  digits = 2,
  caption = "Age of the final sample (N = 138)."
)

kable(
  as.data.frame(table(demo_final_filtered$Gender)),
  col.names = c("Gender", "N"),
  caption = "Gender counts, final sample."
)
```

## 3.3 Comprehension attempts

```{r}
kable(
  tibble(mean_attempts = mean(cleaned_RT_data$attempts_to_pass_four_questions)),
  digits = 2,
  col.names = "Mean attempts to pass the four comprehension questions",
  caption = "Comprehension attempts, final sample."
)
```

## 3.4 Response-time distribution and timeout rate

```{r}
# RT distribution summary: median and 95th percentile from the cleaned
# analysis trials; timeout rate from rt_pre_filter (Section 2.2), i.e.
# BEFORE the 30-s absolute filter — computing it after that filter would
# make it structurally zero (no trial can survive the > 30 s cut and still
# be >= 35 s).
median_val <- median(df_test1$rt, na.rm = TRUE)

p95_val <- quantile(df_test1$rt, 0.95, na.rm = TRUE)

timeout_pct <- mean(rt_pre_filter$rt >= 35, na.rm = TRUE) * 100

kable(
  tibble(
    `Median RT (s)` = median_val,
    `95th percentile RT (s)` = p95_val,
    `Timeout rate (%)` = timeout_pct
  ),
  digits = 2,
  caption = "Response-time distribution (cleaned trials) and timeout rate (pre-filter trials)."
)

# Median RT separately per domain (gains vs losses)
rt_domain_medians <- df_test1 %>%
  group_by(test_part) %>%
  summarise(`Median RT (s)` = median(rt, na.rm = TRUE), .groups = "drop")

kable(rt_domain_medians, digits = 2,
      caption = "Median response time by domain.")
```

## 3.5 Bonus payments

```{r}
# Bonus summary per participant (GBP)
bonus_summary <- data %>%
  filter(!is.na(total_bonus)) %>%  # Keep rows where total_bonus is not NA
  distinct(prolific_pid, total_bonus)

kable(
  tibble(
    `Mean bonus (GBP)` = mean(bonus_summary$total_bonus),
    `Min bonus (GBP)`  = min(bonus_summary$total_bonus),
    `Max bonus (GBP)`  = max(bonus_summary$total_bonus)
  ),
  digits = 2,
  caption = "Bonus payments."
)
```

## 3.6 Completion time and total compensation

```{r}
# Completion time per participant = final cumulative time stamp of their
# session (time_elapsed is cumulative within a session, in ms).
# Total compensation = GBP 10 base payment + total bonus.
completion <- data %>%
  filter(prolific_pid %in% participants_to_keep$prolific_pid) %>%
  group_by(prolific_pid) %>%
  summarise(minutes = max(time_elapsed, na.rm = TRUE) / 1000 / 60,
            .groups = "drop") %>%
  left_join(bonus_summary, by = "prolific_pid") %>%
  mutate(total_compensation = 10 + total_bonus)

kable(
  tibble(
    `Median completion time (min)` = median(completion$minutes),
    `Mean total compensation (GBP)` = mean(completion$total_compensation, na.rm = TRUE),
    `Mean hourly rate (GBP/h)` = mean(completion$total_compensation /
                                        (completion$minutes / 60), na.rm = TRUE)
  ),
  digits = 2,
  caption = "Completion time and compensation (base payment + bonus)."
)
```

# 4. Complex-option choice

Two-sided one-sample t-tests comparing the mean choice proportion of complex
options against the 50% benchmark, separately for the gain and loss conditions,
followed by a two-sided paired t-test comparing the gain and loss conditions.

```{r}
# Per-participant proportion of complex-option choices, by domain
complex_choice_summary_individual <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by(prolific_pid, test_part) %>%
  summarise(
    n_trials = n(),
    n_complex = sum(complex_choice, na.rm = TRUE),
    prop_complex = mean(complex_choice, na.rm = TRUE),
    .groups = "drop"
  )

# Group means with 95% CI
complex_choice_summary_group <-  complex_choice_summary_individual %>%
  group_by(test_part) %>%
  summarise(
    mean_complex_prop = mean(prop_complex, na.rm = TRUE),
    se = sd(prop_complex, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_complex_prop - 1.96 * se,
    upper = mean_complex_prop + 1.96 * se
  )
```

```{r}
# --- One-sample tests vs 0.5 ---
t_gain <- t.test(
  complex_choice_summary_individual$prop_complex[
    complex_choice_summary_individual$test_part == "Gain"
  ],
  mu = 0.5, alternative = "two.sided"
)

t_loss <- t.test(
  complex_choice_summary_individual$prop_complex[
    complex_choice_summary_individual$test_part == "Loss"
  ],
  mu = 0.5, alternative = "two.sided"
)

# --- Paired t-test (Gain vs Loss) ---
complex_wide <- complex_choice_summary_individual %>%
  tidyr::pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = prop_complex
  )

t_gain_vs_loss <- t.test(
  complex_wide$Gain, complex_wide$Loss,
  paired = TRUE, alternative = "two.sided"
)

# --- Tidy results ---
tt_gain <- tidy(t_gain) %>% mutate(test = "Gain vs 0.5")
tt_loss <- tidy(t_loss) %>% mutate(test = "Loss vs 0.5")
tt_gvl  <- tidy(t_gain_vs_loss) %>% mutate(test = "Gain vs Loss (paired)")

# --- Combine ---
t_summary_complex <- bind_rows(tt_gain, tt_loss, tt_gvl) %>%
  select(test, estimate, statistic, parameter, p.value, conf.low, conf.high)

kable(
  t_summary_complex,
  digits = 3,
  caption = "Complex-option choice: one-sample t-tests vs .5 and paired gain-loss t-test."
)
```

```{r}
# --- Bar + jitter plot with significance stars ---
stars_gain <- p_to_stars(t_gain$p.value)
stars_loss <- p_to_stars(t_loss$p.value)
stars_pair <- p_to_stars(t_gain_vs_loss$p.value)

plot_df_points <- complex_choice_summary_individual
plot_df_bars   <- complex_choice_summary_group

# Label domains nicely
plot_df_bars$test_part   <- factor(plot_df_bars$test_part,
                                   levels = c("Gain","Loss"),
                                   labels = c("Gains","Losses"))
plot_df_points$test_part <- factor(plot_df_points$test_part,
                                   levels = c("Gain","Loss"),
                                   labels = c("Gains","Losses"))

# fixed y-positions
y_gain_star <- 0.94
y_loss_star <- 0.94
bracket_y   <- 1.01
label_y     <- 1.05

plot_complex_choice <- ggplot() +
  geom_jitter(
    data = plot_df_points,
    aes(x = test_part, y = prop_complex),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  geom_col(
    data = plot_df_bars,
    aes(x = test_part, y = mean_complex_prop),
    width = 0.55,
    fill = "grey60",
    color = "black",
    alpha = 0.2
  ) +
  geom_errorbar(
    data = plot_df_bars,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.15,
    linewidth = 1.3
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_text(aes(x = "Gains",  y = y_gain_star, label = stars_gain),  vjust = 0, size = 6, fontface = "bold") +
  geom_text(aes(x = "Losses", y = y_loss_star, label = stars_loss), vjust = 0, size = 6, fontface = "bold") +
  geom_segment(aes(x = 1, xend = 1, y = bracket_y - 0.01, yend = bracket_y)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y - 0.01, yend = bracket_y)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y,       yend = bracket_y)) +
  geom_text(aes(x = 1.5, y = label_y, label = stars_pair), vjust = 0, size = 6, fontface = "bold") +
  scale_y_continuous(
    limits = c(0, 1.1),
    breaks = seq(0, 1, by = 0.25),
    expand = c(0, 0)
  ) +
  labs(
    x = "Domain",
    y = "Proportion of complex-option choices",
    title = "Complex-option choices in gain and loss domains"
  ) +
  theme_pub()

plot_complex_choice

ggsave(
  filename = "s1_plot_complex_choice.png",
  plot = plot_complex_choice,
  width = 6,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# 5. Riskier-option choice

Two-sided one-sample t-tests comparing the mean choice proportion of risky
options against the 50% benchmark, separately for the gain and loss conditions,
followed by a two-sided paired t-test comparing the gain and loss conditions.

```{r}
# Per-participant proportion of risky choices ("f" after recoding), by domain
risk_choice_summary <- df_test1 %>%
  group_by(prolific_pid, test_part) %>%
  summarise(
    n_trials = n(),
    n_risky = sum(real_response == "f", na.rm = TRUE),
    prop_risky = mean(real_response == "f", na.rm = TRUE),
    .groups = "drop"
  )

# Group means with 95% CI
risky_choice_summary_group <-  risk_choice_summary %>%
  group_by(test_part) %>%
  summarise(
    mean_risky_prop = mean(prop_risky, na.rm = TRUE),
    se = sd(prop_risky, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_risky_prop - 1.96 * se,
    upper = mean_risky_prop + 1.96 * se
  )
```

```{r}
# --- One-sample (vs .5) and paired t-tests for prop_risky ---
t_risk_gain <- t.test(
  risk_choice_summary$prop_risky[risk_choice_summary$test_part == "Gain"],
  mu = 0.5, alternative = "two.sided"
)

t_risk_loss <- t.test(
  risk_choice_summary$prop_risky[risk_choice_summary$test_part == "Loss"],
  mu = 0.5, alternative = "two.sided"
)


# --- Paired two-sample t-test (Gain vs Loss) ---
# Reshape to wide first
risk_wide <- risk_choice_summary %>%
  tidyr::pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = prop_risky
  )

t_risk_gain_vs_loss <- t.test(
  risk_wide$Gain, risk_wide$Loss,
  paired = TRUE, alternative = "two.sided"
)


# --- Summary table ---
t_summary_risky <- bind_rows(
  tidy(t_risk_gain)        %>% mutate(test = "Gain vs 0.5"),
  tidy(t_risk_loss)        %>% mutate(test = "Loss vs 0.5"),
  tidy(t_risk_gain_vs_loss)%>% mutate(test = "Gain vs Loss (paired)")
) %>%
  select(test, estimate, statistic, parameter, p.value, conf.low, conf.high)

kable(
  t_summary_risky,
  digits = 3,
  caption = "Riskier-option choice: one-sample t-tests vs .5 and paired gain-loss t-test."
)
```

```{r}
stars_gain_r <- p_to_stars(t_risk_gain$p.value)
stars_loss_r <- p_to_stars(t_risk_loss$p.value)
stars_pair_r <- p_to_stars(t_risk_gain_vs_loss$p.value)

plot_pts_r <- risk_choice_summary %>%
  mutate(test_part = factor(test_part,
                            levels = c("Gain","Loss"),
                            labels = c("Gains","Losses")))

plot_bars_r <- risky_choice_summary_group %>%
  mutate(test_part = factor(test_part,
                            levels = c("Gain","Loss"),
                            labels = c("Gains","Losses")))

y_gain_star_r <- 0.8
y_loss_star_r <- 0.8
bracket_y_r   <- 0.87
label_y_r     <- 0.9

plot_risky_choice <- ggplot() +
  geom_jitter(
    data = plot_pts_r,
    aes(x = test_part, y = prop_risky),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  geom_col(
    data = plot_bars_r,
    aes(x = test_part, y = mean_risky_prop),
    width  = 0.55,
    fill   = "grey60",
    color  = "black",
    alpha  = 0.45
  ) +
  geom_errorbar(
    data = plot_bars_r,
    aes(x = test_part, ymin = lower, ymax = upper, color = "#b00000"),
    width = 0.12,
    linewidth = 1.3
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_text(
    aes(x = "Gains",  y = y_gain_star_r, label = stars_gain_r),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  geom_text(
    aes(x = "Losses", y = y_loss_star_r, label = stars_loss_r),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  geom_segment(aes(x = 1, xend = 1, y = bracket_y_r - 0.01, yend = bracket_y_r)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y_r - 0.01, yend = bracket_y_r)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y_r,       yend = bracket_y_r)) +
  geom_text(
    aes(x = 1.5, y = label_y_r, label = stars_pair_r),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.25),
    expand = c(0, 0)
  ) +
  labs(
    x = "Domain",
    y = "Proportion of risky choices",
    title = "Risky choices in gain and loss domains"
  ) +
  theme_pub()

plot_risky_choice

ggsave(
  filename = "s1_plot_risky_choice.png",
  plot = plot_risky_choice,
  width = 6,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# 6. Response times

## 6.1 Median RT difference: complex − simple

```{r}
# ---- per participant × domain × choice (complex vs simple), MEDIAN ONLY ----
# (median RT is the summary reported in the paper and used in the summary
# figure panels)
rt_individual <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  group_by(prolific_pid, test_part, complex_choice) %>%
  summarise(
    median_rt = median(rt, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

# complex_choice: 1 = complex chosen, 0 = simple chosen
rt_wide <- rt_individual %>%
  mutate(choice_label = ifelse(complex_choice == 1, "complex", "simple")) %>%
  select(prolific_pid, test_part, choice_label, median_rt) %>%
  pivot_wider(
    id_cols   = c(prolific_pid, test_part),
    names_from = choice_label,
    values_from = median_rt,
    names_prefix = "median_rt_"
  ) %>%
  # keep only rows with both complex & simple
  filter(!is.na(median_rt_complex), !is.na(median_rt_simple))

# per-participant RT difference (complex − simple)
rt_diff <- rt_wide %>%
  transmute(
    prolific_pid,
    test_part,
    rt_diff = median_rt_complex - median_rt_simple
  ) %>%
  mutate(
    test_part = factor(test_part, levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses"))
  )
```

```{r}
# ---- group summary for bars + 95% CI ----
rt_diff_summary <- rt_diff %>%
  group_by(test_part) %>%
  summarise(
    mean_diff = mean(rt_diff, na.rm = TRUE),
    se        = sd(rt_diff,  na.rm = TRUE) / sqrt(n()),
    lower     = mean_diff - 1.96 * se,
    upper     = mean_diff + 1.96 * se,
    .groups   = "drop"
  )

# ---- tests: one-sample vs 0 (within domain) ----
t_gain_median <- t.test(
  rt_diff$rt_diff[rt_diff$test_part == "Gains"],
  mu = 0
)
t_loss_median <- t.test(
  rt_diff$rt_diff[rt_diff$test_part == "Losses"],
  mu = 0
)

# ---- paired Gain vs Loss (within-subjects) ----
rt_diff_wide <- rt_diff %>%
  pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = rt_diff
  )

t_pair_median <- t.test(
  rt_diff_wide$Gains,
  rt_diff_wide$Losses,
  paired = TRUE
)

# ---- tidy summary table (median only) ----
rt_tests_summary_median <- bind_rows(
  tidy(t_gain_median)  %>% mutate(test = "Gains vs 0",  metric = "Median RT (complex − simple)"),
  tidy(t_loss_median)  %>% mutate(test = "Losses vs 0", metric = "Median RT (complex − simple)"),
  tidy(t_pair_median)  %>% mutate(test = "Gains vs Losses (paired)", metric = "Median RT (complex − simple)")
) %>%
  select(metric, test, estimate, statistic, parameter, p.value, conf.low, conf.high)

kable(
  rt_tests_summary_median,
  digits = 3,
  caption = "Median RT difference (complex − simple): one-sample t-tests vs 0 and paired gain-loss t-test."
)
```

```{r}
stars_gain   <- p_to_stars(t_gain_median$p.value)
stars_loss   <- p_to_stars(t_loss_median$p.value)
stars_pair   <- p_to_stars(t_pair_median$p.value)

# ---- positions for stars and bracket ----
rng       <- range(rt_diff_summary$upper, rt_diff_summary$lower, na.rm = TRUE)
span      <- diff(rng)
span      <- ifelse(span == 0, 0.1, span)  # fallback

y_gain_star <- 4.2
y_loss_star <- 4.2

bracket_y   <- 4.2
label_y     <- 5.3

# ---- final plot object ----
plot_rt_median_diff <- ggplot() +
  # individual participant differences
  geom_jitter(
    data  = rt_diff,
    aes(x = test_part, y = rt_diff),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  # group mean bars (semi-transparent)
  geom_col(
    data  = rt_diff_summary,
    aes(x = test_part, y = mean_diff),
    width = 0.55,
    fill  = "grey60",
    color = "black",
    alpha = 0.45
  ) +
  # 95% CI
  geom_errorbar(
    data = rt_diff_summary,
    aes(x = test_part, ymin = lower, ymax = upper),
    width = 0.12,
    linewidth = 1
  ) +
  # zero reference line
  geom_hline(yintercept = 0, linetype = "dashed") +
  # stars for one-sample tests
  geom_text(
    aes(x = "Gains",  y = y_gain_star, label = stars_gain),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  geom_text(
    aes(x = "Losses", y = y_loss_star, label = stars_loss),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  # bracket + star for Gain vs Loss
  geom_segment(aes(x = 1, xend = 1, y = 5, yend = 5.2)) +
  geom_segment(aes(x = 2, xend = 2, y = 5, yend = 5.2)) +
  geom_segment(aes(x = 1, xend = 2, y = 5.2, yend = 5.2)) +
  geom_text(
    aes(x = 1.5, y = label_y, label = stars_pair),
    vjust = 0, size = 6, fontface = "bold"
  ) +
  labs(
    x = "Domain",
    y = "Median RT difference (complex − simple, s)",
    title = "Median RT differences for complex vs simple choices"
  ) +
  coord_cartesian(
    ylim = c(-3, 7)
  ) +
  theme_pub()

plot_rt_median_diff


ggsave(
  filename = "s1_plot_rt_median_diff.png",
  plot = plot_rt_median_diff,
  width = 6,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

## 6.2 Median RT difference: riskier − safer

```{r}
# Recode choice: risky = 1, safe = 0
df_rt_risky <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  mutate(
    risky_choice = ifelse(real_response == "f", 1,
                          ifelse(real_response == "j", 0, NA))
  ) %>%
  filter(!is.na(risky_choice))

# ---- summarise median RT per participant × domain × risky/safe ----
rt_individual_risky <- df_rt_risky %>%
  group_by(prolific_pid, test_part, risky_choice) %>%
  summarise(
    median_rt = median(rt, na.rm = TRUE),
    .groups = "drop"
  )

# ---- wide format (median_rt_risky vs median_rt_safe) ----
rt_wide_risky <- rt_individual_risky %>%
  mutate(choice_label = ifelse(risky_choice == 1, "risky", "safe")) %>%
  select(prolific_pid, test_part, choice_label, median_rt) %>%
  pivot_wider(
    id_cols = c(prolific_pid, test_part),
    names_from = choice_label,
    values_from = median_rt,
    names_prefix = "median_rt_"
  ) %>%
  filter(!is.na(median_rt_risky), !is.na(median_rt_safe))

# ---- compute difference (risky − safe) ----
rt_diff_risky <- rt_wide_risky %>%
  transmute(
    prolific_pid,
    test_part = factor(test_part,
                       levels = c("Gain", "Loss"),
                       labels = c("Gains", "Losses")),
    rt_diff = median_rt_risky - median_rt_safe
  )
```

```{r}
# ---- one-sample t-tests vs 0 ----
t_gain_risky <- t.test(
  rt_diff_risky$rt_diff[rt_diff_risky$test_part == "Gains"],
  mu = 0
)

t_loss_risky <- t.test(
  rt_diff_risky$rt_diff[rt_diff_risky$test_part == "Losses"],
  mu = 0
)

# ---- paired t-test between domains ----
rt_diff_risky_wide <- rt_diff_risky %>%
  pivot_wider(
    id_cols = prolific_pid,
    names_from = test_part,
    values_from = rt_diff
  )

t_pair_risky <- t.test(
  rt_diff_risky_wide$Gains,
  rt_diff_risky_wide$Losses,
  paired = TRUE
)

# ---- tidy table ----
rt_tests_summary_risky <- bind_rows(
  tidy(t_gain_risky) %>% mutate(test = "Gains vs 0"),
  tidy(t_loss_risky) %>% mutate(test = "Losses vs 0"),
  tidy(t_pair_risky) %>% mutate(test = "Gains vs Losses (paired)")
) %>%
  mutate(metric = "Median RT (risky − safe)") %>%
  select(metric, test, estimate, statistic, parameter, p.value, conf.low, conf.high)

kable(
  rt_tests_summary_risky,
  digits = 3,
  caption = "Median RT difference (riskier − safer): one-sample t-tests vs 0 and paired gain-loss t-test."
)
```

```{r}
# ---- summary for bars ----
rt_diff_risky_summary <- rt_diff_risky %>%
  group_by(test_part) %>%
  summarise(
    mean_diff = mean(rt_diff, na.rm = TRUE),
    se        = sd(rt_diff, na.rm = TRUE) / sqrt(n()),
    lower     = mean_diff - 1.96 * se,
    upper     = mean_diff + 1.96 * se,
    .groups = "drop"
  )

# ---- stars ----
stars_gain <- p_to_stars(t_gain_risky$p.value)
stars_loss <- p_to_stars(t_loss_risky$p.value)
stars_pair <- p_to_stars(t_pair_risky$p.value)

# vertical spacing
rng  <- range(rt_diff_risky_summary$lower, rt_diff_risky_summary$upper, na.rm = TRUE)
span <- diff(rng); if (span == 0) span <- 0.1

y_gain_star <- 2.1
y_loss_star <- 2.1

bracket_y <- 2.7
label_y   <- 2.9

# ---- plot ----
plot_rt_risky <- ggplot() +
  geom_jitter(
    data = rt_diff_risky,
    aes(x = test_part, y = rt_diff),
    width = 0.10, height = 0,
    alpha = 0.45, size = 2
  ) +
  geom_col(
    data = rt_diff_risky_summary,
    aes(x = test_part, y = mean_diff),
    width = 0.55,
    fill  = "grey60",
    color = "black",
    alpha = 0.45
  ) +
  geom_errorbar(
    data = rt_diff_risky_summary,
    aes(x = test_part, ymin = lower, ymax = upper),
    width = 0.12,
    linewidth = 1
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +

  geom_text(aes(x = "Gains",  y = y_gain_star, label = stars_gain),
            vjust = 0, size = 6, fontface = "bold") +
  geom_text(aes(x = "Losses", y = y_loss_star, label = stars_loss),
            vjust = 0, size = 6, fontface = "bold") +

  geom_text(aes(x = 1.5, y = y_loss_star+0.6, label = stars_pair),
            vjust = 0, size = 6, fontface = "bold") +

  geom_segment(aes(x = 1, xend = 1, y = bracket_y - 0.01*span, yend = bracket_y)) +
  geom_segment(aes(x = 2, xend = 2, y = bracket_y - 0.01*span, yend = bracket_y)) +
  geom_segment(aes(x = 1, xend = 2, y = bracket_y,               yend = bracket_y)) +

  labs(
    x = "Domain",
    y = "Median RT difference (risky − safe, s)",
    title = "Median RT differences for risky vs. safe choices"
  ) +
  coord_cartesian(
    ylim = c(-3,3)
  ) +
  theme_pub()

plot_rt_risky

ggsave(
  filename = "s1_plot_rt_risky.png",
  plot = plot_rt_risky,
  width = 6,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# 7. Bayesian logistic mixed-effects regression

```{r}
# Trial-level moments of each option (mean, SD, skewness) and their difference
df_test1 <- df_test1 %>%
  rowwise() %>%
  mutate(
    # --- Option A ---
    mean_A = probA1 * payA1 + probA2 * payA2,
    var_A  = probA1 * (payA1 - mean_A)^2 + probA2 * (payA2 - mean_A)^2,
    sd_A   = sqrt(var_A),
    skew_A = ifelse(sd_A > 0,
                    (probA1 * (payA1 - mean_A)^3 + probA2 * (payA2 - mean_A)^3) / (sd_A^3),
                    0),

    # --- Option B ---
    mean_B = probB1 * payB1 + probB2 * payB2,
    var_B  = probB1 * (payB1 - mean_B)^2 + probB2 * (payB2 - mean_B)^2,
    sd_B   = sqrt(var_B),
    skew_B = ifelse(sd_B > 0,
                    (probB1 * (payB1 - mean_B)^3 + probB2 * (payB2 - mean_B)^3) / (sd_B^3),
                    0),

    # --- Skewness difference ---
    skew_diff = skew_A - skew_B
  ) %>%
  ungroup()
```

```{r}
# Regression predictors: response coded relative to the left option,
# side-corrected (via risk_index) and z-scored EV / SD / skewness differences
df_test1_fix <- df_test1 %>%
  mutate(
    real_response_num = ifelse(response == "f", 1, 0),
    test_part      = factor(test_part, levels = c("Gain","Loss")),  # domain
    complexity_left = ifelse(risk_index == 1, ifelse(complexity_option == "A", 1L, 0L), ifelse(complexity_option == "A", 0L, 1L)),
    SD_diff_numeric = as.numeric(SD_diff_level)*as.numeric(risk_index),
    EVD_numeric     = as.numeric(EV_diff_level)*as.numeric(risk_index),
    EVD_z = scale(EVD_numeric, center = TRUE, scale = TRUE)[,1],
    SD_diff_z = scale(SD_diff_numeric, center = TRUE, scale = TRUE)[,1],
    skew_diff_z = scale(skew_diff*as.numeric(risk_index), center = TRUE, scale = TRUE)[,1]
  )
```

```{r}
library(brms)

# Long-running fit (4 chains x 4000 iterations; expect roughly 20-60 min on
# a first run). The fitted object is cached to m_brm_study1.rds and re-used
# on subsequent runs; caching only stores/reloads the fitted object and does
# not affect the estimates in any way. Delete the .rds to force a refit.
if (file.exists("m_brm_study1.rds")) {
  m_all_rs_brm <- readRDS("m_brm_study1.rds")
} else {
  m_all_rs_brm <- brm(
    real_response_num ~
      EVD_z + skew_diff_z +
      complexity_left * SD_diff_z * test_part +
      (1 | prolific_pid),
    data = df_test1_fix,
    family = bernoulli(link = "logit"),
    chains = 4, cores = 4, iter = 4000,
    control = list(adapt_delta = 0.95, max_treedepth = 12)
  )
  saveRDS(m_all_rs_brm, "m_brm_study1.rds")
}

summary(m_all_rs_brm)
```

```{r}
# Parameter table used for the paper's regression Table
library(bayestestR)
library(parameters)

# effects = "all" also reports the random-effect SDs (the model includes
# random slopes for complexity, risk, and domain, not only random intercepts)
table_results <- model_parameters(m_all_rs_brm, ci = 0.95, effects = "all")
print_md(table_results)
```

# 8. EV-dominant (catch) trial accuracy

```{r}
# Catch trials: one option dominates by expected value (|EV difference| = 20).
# Recode responses for screen side, then score whether the dominant option
# was chosen.
df_test2 <- cleaned_RT_data %>%
  filter(trial_type_label == "test" & skew == "catch") %>%
  mutate(
    EV_diff_level = as.numeric(EV_diff_level)
    )


df_test2 <- df_test2 %>%
  mutate(real_response = case_when(
    risk_index == 1 & response == "f" ~ "f",
    risk_index == 1 & response == "j" ~ "j",
    risk_index == -1 & response == "f" ~ "j",
    risk_index == -1 & response == "j" ~ "f"
  )) %>%
  mutate(good_decision = case_when(
    EV_diff_level == 20  & real_response == "f" ~ 1,
    EV_diff_level == -20 & real_response == "j" ~ 1,
    TRUE ~ 0
  ))
```

```{r}
# Per-participant catch-trial accuracy and group means with 95% CI
catch_accuracy_individual <- df_test2 %>%
  group_by(prolific_pid, test_part) %>%
  summarise(
    n_trials = n(),
    prop_catch_accuracy = mean(good_decision),
    .groups = "drop"
  )



catch_accuracy_group <-  catch_accuracy_individual %>%
  group_by(test_part) %>%
  summarise(
    mean_catch_prop = mean(prop_catch_accuracy, na.rm = TRUE),
    se = sd(prop_catch_accuracy, na.rm = TRUE) / sqrt(n())
  ) %>%
  mutate(
    lower = mean_catch_prop - 1.96 * se,
    upper = mean_catch_prop + 1.96 * se
  )

kable(
  catch_accuracy_group,
  digits = 3,
  caption = "EV-dominant (catch) trial accuracy: group means with 95% CI."
)
```

```{r}
# --- One-sample (vs .5) and paired two-sample t-tests ---
t_catch_gain <- t.test(
  catch_accuracy_individual$prop_catch_accuracy[catch_accuracy_individual$test_part == "Gain"],
  mu = 0.5, alternative = "two.sided"
)

t_catch_loss <- t.test(
  catch_accuracy_individual$prop_catch_accuracy[catch_accuracy_individual$test_part == "Loss"],
  mu = 0.5, alternative = "two.sided"
)

# Paired: same participants in Gain & Loss
catch_wide <- catch_accuracy_individual %>%
  pivot_wider(id_cols = prolific_pid,
              names_from = test_part,
              values_from = prop_catch_accuracy)

t_catch_gain_vs_loss <- t.test(
  catch_wide$Gain, catch_wide$Loss,
  paired = TRUE, alternative = "two.sided"
)

# --- Summary table ---
t_summary_catch <- bind_rows(
  tidy(t_catch_gain)        %>% mutate(test = "Gain vs 0.5"),
  tidy(t_catch_loss)        %>% mutate(test = "Loss vs 0.5"),
  tidy(t_catch_gain_vs_loss)%>% mutate(test = "Gain vs Loss (paired)")
) %>%
  select(test, estimate, statistic, parameter, p.value, conf.low, conf.high)

kable(
  t_summary_catch,
  digits = 3,
  caption = "EV-dominant (catch) trial accuracy: one-sample t-tests vs .5 and paired gain-loss t-test."
)
```

```{r}
stars_gain   <- p_to_stars(t_catch_gain$p.value)
stars_loss   <- p_to_stars(t_catch_loss$p.value)
stars_pair   <- p_to_stars(t_catch_gain_vs_loss$p.value)


# --- prep for plotting: rename x-axis to Domain (Gains / Losses) ---
plot_pts <- catch_accuracy_individual %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  mutate(domain = factor(ifelse(test_part == "Gain", "Gains", "Losses"),
                         levels = c("Gains", "Losses")))

plot_bars <- catch_accuracy_group %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  mutate(domain = factor(ifelse(test_part == "Gain", "Gains", "Losses"),
                         levels = c("Gains", "Losses")))

# vertical positions for stars & bracket
y_max    <- max(plot_bars$upper, na.rm = TRUE)
pad      <- 0.03
y_gain_s <- 0.8
y_loss_s <- 0.8
br_y     <- y_max + 0.06
lab_y    <- br_y + 0.03
pair_lab <- paste0(stars_pair)

plot_catch_acc <- ggplot() +
  # individual participants
  geom_jitter(
    data = plot_pts,
    aes(x = domain, y = prop_catch_accuracy),
    width = 0.12, height = 0,
    alpha = 0.45, size = 2
  ) +
  # group means
  geom_col(
    data = plot_bars,
    aes(x = domain, y = mean_catch_prop),
    width  = 0.6,
    fill   = "grey80",
    color  = "black",
    alpha  = 0.9
  ) +
  # 95% CI bars
  geom_errorbar(
    data = plot_bars,
    aes(x = domain, ymin = lower, ymax = upper),
    width     = 0.12,
    linewidth = 1
  ) +
  # chance line
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  # one-sample stars above bars
  geom_text(aes(x = "Gains",  y = y_gain_s, label = stars_gain),  vjust = 0, size = 5) +
  geom_text(aes(x = "Losses", y = y_loss_s, label = stars_loss), vjust = 0, size = 5) +
  # paired bracket and label
  geom_segment(aes(x = 1, xend = 1, y = br_y - 0.01, yend = br_y)) +
  geom_segment(aes(x = 2, xend = 2, y = br_y - 0.01, yend = br_y)) +
  geom_segment(aes(x = 1, xend = 2, y = br_y,        yend = br_y)) +
  geom_text(aes(x = 1.5, y = br_y+0.1, label = pair_lab), size = 8)+
  coord_cartesian(ylim = c(0, 1)) +
  labs(
    x = "Domain",
    y = "EV dominant trial accuracy (proportion correct)",
    title = "EV dominant trial accuracy in gains and losses"
  ) +
  theme_pub(base_size = 14)

plot_catch_acc

ggsave(
  filename = "s1_plot_catch_acc.png",
  plot = plot_catch_acc,
  width = 6,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# 9. Cognitive-ability correlations

```{r}
# Cognitive-ability score per participant: mean of Berlin Numeracy Test (BNT)
# and Health/Money Test (HMT) accuracies
result_ca <- data %>%
  group_by(prolific_pid) %>%
  dplyr::summarize(
    accuracy_BNT = mean(accuracy_BNT, na.rm = TRUE),
    accuracy_HMT = mean(accuracy_HMT, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(accuracy_BNT, accuracy_HMT), names_to = "accuracy_type", values_to = "score")

result_ca <- result_ca %>%
  group_by(prolific_pid) %>%
  dplyr::summarize(ca_results = mean(score, na.rm = TRUE))
```

## 9.1 CA × complex-option choice

```{r}
# Spearman correlation between cognitive ability and complex-choice proportion,
# per domain, with plots

# --- join data ---
joined_df <- inner_join(result_ca, complex_choice_summary_individual,
                        by = "prolific_pid") %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  mutate(test_part = factor(test_part, levels = c("Gain", "Loss")))

# --- Spearman correlation per condition (Gain / Loss) ---
corr_results <- joined_df %>%
  group_by(test_part) %>%
  summarise(
    cor_test = list(tidy(cor.test(ca_results, prop_complex, method = "spearman"))),
    .groups = "drop"
  ) %>%
  unnest(cor_test)

kable(
  corr_results,
  digits = 3,
  caption = "Spearman correlations between cognitive ability and complex-option choice proportion, per domain (Loss row: Appendix A)."
)

# =========================================================
# FUNCTION: build one plot for Gain or Loss
# =========================================================
make_plot_ca_corr_domain <- function(domain_label, df, corr_df){

  df_domain <- df %>% filter(test_part == domain_label)

  label_df <- corr_df %>%
    filter(test_part == domain_label) %>%
    mutate(
      stars = p_to_stars(p.value),
      label = paste0("ρ = ", sprintf("%.2f", estimate), "  ", stars)
    )

  ggplot(df_domain, aes(x = ca_results, y = prop_complex)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
    geom_text(
      data = label_df,
      aes(x = -Inf, y = Inf, label = label),
      hjust = -0.05, vjust = 1.1,
      size = 7, fontface = "bold",
      inherit.aes = FALSE
    ) +
    labs(
      x = "CA performance",
      y = "Proportion of complex choices",
      title = paste0("Correlation between CA and complex choice (", domain_label, ")")
    ) +
    theme_pub(base_size = 14)
}

# Gain-domain panel (main text)
plot_ca_corr_gain <- make_plot_ca_corr_domain("Gain", joined_df, corr_results)
plot_ca_corr_gain

ggsave(
  filename = "s1_plot_ca_corr_gain.png",
  plot = plot_ca_corr_gain,
  width = 6.5,
  height = 5,
  units = "in",
  dpi = 600
)
```

### Appendix A: loss-domain panel

```{r}
# Loss-domain version of the CA x complex-choice correlation plot; reported
# in Appendix A of the manuscript.
plot_ca_corr_loss <- make_plot_ca_corr_domain("Loss", joined_df, corr_results)
plot_ca_corr_loss

ggsave(
  filename = "s1_plot_ca_corr_loss.png",
  plot = plot_ca_corr_loss,
  width = 6.5,
  height = 5,
  units = "in",
  dpi = 600
)
```

## 9.2 CA × EV-consistent choice

```{r}
# Cognitive ability vs EV-consistent choice

# --- 1. Compute EV-consistent choice per trial ---
df_ev <- df_test1 %>%
  filter(test_part %in% c("Gain", "Loss")) %>%
  filter(EV_diff_level != 0) %>%              # ignore EV=0 trials
  mutate(
    domain = factor(ifelse(test_part == "Gain", "Gains", "Losses"),
                    levels = c("Gains", "Losses")),

    # which option has higher EV?
    higher_EV_option = ifelse(EV_diff_level > 0, "A", "B"),

    # chosen option (f = left/A, j = right/B)
    chosen_option = ifelse(real_response == "f", "A", "B"),

    # EV-consistent = chose higher EV option
    ev_consistent = ifelse(chosen_option == higher_EV_option, 1, 0)
  )


# --- 2. Proportion per participant × domain ---
ev_consistency_individual <- df_ev %>%
  group_by(prolific_pid, domain) %>%
  summarise(
    prop_ev_consistent = mean(ev_consistent, na.rm = TRUE),
    .groups = "drop"
  )


# --- 3. Join with CA scores ---
joined_ev <- inner_join(result_ca, ev_consistency_individual, by = "prolific_pid")


# --- 4. Spearman correlations separately for Gains and Losses ---
corr_ev_results <- joined_ev %>%
  group_by(domain) %>%
  summarise(
    cor_test = list(tidy(cor.test(ca_results,
                                  prop_ev_consistent,
                                  method = "spearman"))),
    .groups = "drop"
  ) %>%
  unnest(cor_test)

kable(
  corr_ev_results,
  digits = 3,
  caption = "Spearman correlations between cognitive ability and EV-consistent choice proportion (zero-EV-difference trials excluded)."
)
```

```{r}
# add stars and clean label formatting
label_df_ev <- corr_ev_results %>%
  mutate(
    stars = p_to_stars(p.value),
    label = paste0("ρ = ",
                   sprintf("%.2f", estimate),
                   "  ", stars)
  )

# --- APA-style plot with stars ---
plot_ev_corr <- ggplot(joined_ev,
                       aes(x = ca_results, y = prop_ev_consistent)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = T, linewidth = 0.8) +
  facet_wrap(~ domain) +
  geom_text(data = label_df_ev,
            aes(label = label),
            x = -Inf, y = Inf,
            hjust = -0.05, vjust = 1.1,
            size = 7,
            fontface = "bold") +
  labs(
    x = "CA performance",
    y = "Proportion of EV-consistent choices",
    title = "Correlation between CA and EV-consistent choice proportion"
  ) +
  theme_pub(base_size = 14)

plot_ev_corr

ggsave(
  filename = "s1_plot_ev_corr.png",
  plot = plot_ev_corr,
  width = 7.5,           # in inches
  height = 5,
  dpi = 600            # high resolution
)
```

# 10. Manuscript summary figures

```{r}
# Exploratory summary figure: CA correlation (gain), catch accuracy,
# CA x EV-consistency
p1 <- ggdraw() + draw_image("s1_plot_ca_corr_gain.png",  scale = 0.98)
p2 <- ggdraw() + draw_image("s1_plot_ev_corr.png",  scale = 0.98)
p3 <- ggdraw() + draw_image("s1_plot_catch_acc.png", scale = 0.98)

# top row: two correlation panels
top_row <- plot_grid(
  p1, p3,
  labels = c("(a)", "(b)"),
  label_size = 16,
  label_fontface = "bold",
  ncol = 2,
  align = "hv"
)

# bottom row: catch accuracy spanning full width
bottom_row <- plot_grid(
  p2,
  labels = "(c)",
  label_size = 16,
  label_fontface = "bold"
)

# combine rows into final figure
summary_explore_analysis <- plot_grid(
  top_row,
  bottom_row,
  ncol = 1,
  rel_heights = c(1, 0.9)   # tweak if you want more space for bottom plot
)

ggsave(
  "summary_explore_analysis.png",
  summary_explore_analysis,
  width  = 10,
  height = 9,
  dpi    = 600
)
```

```{r}
# Main 2x2 summary figure: choice proportions and RT contrasts
p1 <- ggdraw() + draw_image("s1_plot_complex_choice.png", scale = 1)
p2 <- ggdraw() + draw_image("s1_plot_risky_choice.png", scale = 1)
p3 <- ggdraw() + draw_image("s1_plot_rt_median_diff.png", scale = 1)
p4 <- ggdraw() + draw_image("s1_plot_rt_risky.png", scale = 1)

main_summary_plot <- plot_grid(
  p1, p2,
  p3, p4,
  labels = c("(a)", "(b)", "(c)", "(d)"),
  label_size = 16,
  label_fontface = "bold",
  ncol = 2,
  align = "hv",
  rel_heights = c(1.05, 1)
)

ggsave(
  "summary_main_analysis.png",
  main_summary_plot,
  width = 13,
  height = 10,
  dpi = 600
)
```
